---
title: "Working with diatom data in the neotoma2 R package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with diatom data in the neotoma2 R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(neotoma2)
library(neotoma)
```

In this document some code was borrowed from the Neotoma2 markdown document prepared by Simon Goring and Socorro Dominguez and the Neotoma original package. 

In chunks, use "neotoma::get_site(whatever)" when using old neotoma package with neotoma2 

One of the main goals of Neotoma is to pull records out of Neotoma into the R workflow.  

# Introduction

The ultimate purpose of this document is to provide general guidance and specific code examples for those wanting to use Neotoma R2 to retrieve and work with Neotoma diatom data.  An initial purpose is to help programmers (Simon, Socorro) develop Neotoma2 resources by providing a structured description of how diatom researchers may wish to use N-R2, and specific tasks they may wish to accomplish.

This document is organized to address two main goals of users, to browse the database to identify datasets that meet specified criteria, and to create output files that can be used for further analysis.  Other goals are to create plots and maps and to analyze data using R code.


# Browse data

The browse section describes ways of searching for available diatom datasets using a variety of criteria, either alone or in combination.  Criteria include those available in Neotoma Explorer, plus several more. Examples also include options for viewing and saving output results.

## Find diatom datasets using one or more search criteria.

General task specification format: “Find all diatom datasets that meet the following criteria.”  [Chunks below show generic code specifying how to search / filter by a single criterion, followed by an example] 

### [Generic format of a complete chunk to browse using a single criterion]

```{R genericGetDataset, eval = FALSE}
diat_1 <- get_datasets(argument_1, argument_2)
```

Example for Dataset

```{R retrieveDiatoms, eval = FALSE}
NE_US_DTM<-get_dataset(datasettype='diatom', loc =c(-80,40,-65,48))
```


### Geographic region

```{r getGeoPolUnits,eval=FALSE}
gpids <- neotoma::get_table(table.name='GeoPoliticalUnits')
```

Example - get sites from multiple geopolitical units (e.g., states)
```{r retrieveByGeopol}
library(dplyr)
library(purrr)

gpids <- c(8412, 7990, 7934, 7326, 7923, 8981, 6442, 7368)
ne_sites <- map(gpids, function(x)
  neotoma:::get_site(gpid = x)) %>%
  neotoma::bind()
```

####  Geopolitical units (e.g., states)
```{R callDatasetsbygpid,eval=FALSE}
LOC <- neotoma2::get_table(x='geopoliticalunits', limit = 10000)
diat_geo <- get_dataset(gpid = X)
```

Example - get_dataset for an individual state
```{R gpidExamplePenn}
LOC <- get_table(table.name='GeoPoliticalUnits')
Geo_PA <- get_dataset(gpid = "Pennsylvania")
GEO_PA1 <- get_dataset(gpid = 8412 )
```

#### Latitude and Longitude boxes

```{R diatomsbyBBox}
diat_LL <- get_datasets(datasettype="diatom", loc=BOUNDING BOX)
```

Example of get_datasets by location with bounding box of lat & longs (West/South/East/North)

```{R diatomsbyBBoxExecuted}
diat_NE <- get_datasets(datasettype="diatom",loc =c(-80,40,-65,48)) 
```

#### Elevation (m)

```{R sitesbyElev,eval=FALSE}
diat_el <- get_sites(`altmin()`, `altmax()`)
```
Example of getting datasets for sites between minimum and maximum elevations 
```{R getSitesByElevExample}
diat_el <- get_sites("altmin(5000)","altmax(5500)") #Will need to look over at a later time 
diat_el2 <- get_datasets(datasetid= diat_el)
```

### Dataset type - Diatom, Diatom Surface Sample, Diatom Top Bottom

```{r getDiatomsDSType, eval=FALSE}
neDiatoms <- get_datasets(datasettype="diatom surface samples", loc=BOUNDING BOX)

neDiatoms <- get_datasets(dois=c(...)) # DOI's
```

Example of combining datatypes and using the variable to get dataset 
```{r getDiatoms, eval=FALSE} 
# Get datasets by individual diatom datatype 
neDiatoms_ss <- get_datasets(datasettype="diatom surface samples", loc=BOUNDING BOX)
neDiatoms_tb <- get_datasets(datasettype="diatom top-bottom", loc=BOUNDING BOX)
neDiatoms_dt <- get_datasets(datasettype="diatom", loc=BOUNDING BOX)
# Get datasets by multiple diatom datatypes 
neDiatoms_sstb <- combine(neDiatoms_ss,neDiatoms_tb,neDiatoms_dt) 
neDiatoms_com <- get_datasets(neDiatoms_sstb)
```

Example of getting surface sample through filtering 
```{R datasetByTaxa,eval=FALSE}
## Simon Edited this:
diatomss<- get_dataset(datasettype = 'diatom surface sample')
diatom<- get_dataset(datasettype = 'diatom')
#Sapplyy function is used to return diatomss and with the function it is reading the metadata and selectind dataset,id to query from first 20 lakes 
subsetss <- sapply(diatomss, 
                 function(x) {
                   x$dataset.meta$dataset.id
                   })[1:20]

subset <- sapply(diatom, 
                 function(x) {
                   x$dataset.meta$dataset.id
                   })[1:20]

if('filename' %in% list.files()) {
  diat_dl <- readRDS('filename') # It returns the file that is used 
} else {
  diat_dl <- get_download(subset)
  diat_dlss <- get_download(subsetss)
  diat_dl <- bind(diat_dl, diat_dlss)
  saveRDS(diat_dl, 'filename') # It saves this as an object of the download with file
}

wideCounts <- compile_downloads(diat_dl) 
#neotoma function converts multiple downloads into single object 
minRow <- wideCounts %>% 
  group_by(dataset) %>% #dataset was returned from the table 
  summarise(min(depth)) #dplyr function to create a dataframe 

wideCounts %>% filter(!is.na(Eunotia.diodon))
#dplyr function used to subset a data frame
diat_taxa <- taxa(diat.dl)
```

### DOI's
```{r getDiatomsDSType, eval=FALSE}
neDiatoms <- get_datasets(datasettype="diatom surface samples", loc=BOUNDING BOX)

neDiatoms <- get_datasets(dois=c(...)) # DOI's
```


#### Site - Name(s) or ID(s)
```{r get_site,eval=FALSE}
diat_1<-get_site(gpid = x)
diat_2<-get_site(sitename= x)
diat_3<-get_site(siteid= x)
```
Examples of getting sites by geopolitical units, sitenames, and siteid
```{R SitebyName}
diat_1<-get_site(gpid = 6198) # Geopolitical ID for Alaska from Table
diat_2<-get_site(sitename= 'Schrader Pond')
diat_3<-get_site(siteid= 27408)
```


#### Collection type
```{R Diatom Collection Type,eval=FALSE}
get_dataset(CollType='x') # Unsure
```
Example of getting dataset by collection id
```{R sitesByCollType}
get_dataset(CollType='Core','Section')
```

#### Constituent database
```{R SitesbyConDB,eval=FALSE}
Browse('x')
```

Example of finding constituent databases

```{R SitesbyConDBExample}
Browse('Database') # Not actual code but browse seems like a right choice
```

#### Sample keyword
```{R SitesbyKW, eval=FALSE}
get_dataset(search('x'))
```

Example of using the search function returning keywords 

```{R}
get_dataset(search('keywords'))# Search could find so maybe
```

#### Within or intersecting a specified time interval 
```{R Diatom Time Interval,eval=FALSE}

```
Example
```{R}

```

#### Member of one or more aggregate datasets
```{R Diatom Aggregate Dataset,eval=FALSE}
library(neotoma)
library(neotoma2)


```
Example
```{R}

```

#### Related to one or more publications
```{R Diatom Publication,eval=FALSE}
diat_pub <- get_datasets(search())
```
Example
```{R}
diat_pub <- get_publication()
```

#### From sites with specified range of water chemistry characteristics
```{r getWaterChem, eval=FALSE}
neDiatoms <- get_datasets(datasettype="diatom surface samples", loc=BOUNDING BOX)
```
Example
```{R}

```


#### From sites where one or more specified water chemistry characteristics were measured
```{R Diatom Multiple Water Chemistry,eval=FALSE}

```
Example
```{R}

```

#### Diatom Sample analyst name
```{R Diatom Sample Analyst,eval=FALSE}
<- get_contact()
```
Example
```{R}

```

#### Investigator name / person name
```{R Diatom Investigator,eval=FALSE}
Contact <- get_contact(contactname = 'X')

```
Example of returning a contact id given the contact name
```{R}
Don <- get_contact(contactname = 'Donald F. Charles')
```

#### Diatom counts include one or more specified diatom taxa

```{R datasetByTaxa,eval=FALSE}
diatoms<- get_dataset(datasettype = 'diatoms')
diat_dl <- get_download(diatoms)
diat_taxa <- taxa(diat.dl)
```

Example of getting  the downloads of a datasettype 

```{R datasetByTaxaExample}
diatoms<- get_dataset(datasettype = 'diatoms')
diat_dl <- get_download(diatoms)
diat_taxa <- taxa(diat.dl)
```

#### Diatom counts include only taxa with abundance greater than x percent (in sample or core)

```{R Diatom Counts,eval=FALSE}

```

Example

```{R}

```


#### Site area larger or smaller than specified size

```{R Diatom Site Area,eval=FALSE}

```

Example

```{R}

```

#### Depository in which diatom slides are located

```{R Depository,eval=FALSE}
diat.data <- get_dataset(datasettype='diatom') # access to neotoma explorer but needs to specify 
browse(diat.data)
```

Example

```{R}

```

#### Chronology (y/n) and characteristics
```{R Chronology,eval=FALSE}
<-get_chroncontrol(x) # Need to check if this relates to chronology or bacon 
```

Example
```{R ChronologyExample}

```


### Format for browsing using multiple criteria.  Examples: [could get complicated]  General rules and tips.  [Add as they become known]

#### Find all diatom datasets of any datatype

#### Find all diatom datasets of any datatype within a geographic region

#### Find all diatom surface sample datasets within a geographic region

#### Find datasets .... with surface area larger than x ha .......

### Other options / Misc:

#### Is water chemistry available?  (List variables included) Is chronology available?  Upper and lower age bounds.  Aggregate dataset(s) of which the dataset is a member.

## Create output of browse results.  Minimum “list” of dataset, collection, and / analysis unit ids for further browsing and use as input to other R functions. 
 
Standard comprehensive output file for use in detailed browsing. Variables should include all of the following (variables are grouped by their location on Tilia tabs):

### Site info:
#### Neotoma Site ID, Site name, Latitude, Longitude, Latitude width, Longitude width, Country, State, County, Admin Unit (3rd Geographic Unit), Lake characteristics

### Collection unit:
#### Collection unit type, Collection unit name, Collection unit handle, Location in site, Collection device, Date collected, Water depth of collection

### Dataset:
#### Dataset type - (Diatom, Diatom surface sample, or Diatom top-bottom), Primary publication, Investigators [or first author of publication], Sample analyst

```{R,eval=FALSE}

```

Example
```{R}

```

# Create data output files  

The create data output files section -  standard formats for diatom counts, metadata, chronology, water chemistry.  It provides examples for retrieving data and creating output files in a few basic formats.  Formats include those that can be used as input to data analysis programs - R packages and others.  In general, they should include all related data; the amount of data can be reduced later.They will include several descriptive variables and be easily readable. [input to code chunks can include output from browse tasks]

## Basic files / matrices

```{R,eval=FALSE}

```

Example
```{R}

```

### Diatom counts - single sites and multiple sites; aggregate data set; stratigraphic vs surface sample vs top-bottom; percentages vs raw counts.

```{R,eval=FALSE}

```
Example of a single site with getting count data  

```{r get_singleSitebyName, eval=FALSE}
schrader <- get_site('Schrader Pond') # % can be used to approximate sitenames
schrader.data <- get_dataset(schrader)
schrader.dl <- get_download(schrader.data)
schrader.cnt <- counts(schrader.dl)
sapply(schrader.cnt, dim) 
print(schrader.dl)
```


```{R}
library (neotoma)
# Find Neotoma Site Id for Upper Wallface Pond 
uwallface <- get_site (sitename = "Upper Wallface%")
# Find datasets and IDs associated with U Wallface Pond
uwallface.dl<- get_dataset(uwallface)  # Review datasets available note ID of the one of interest
# download data for diatom core dataset (Dataset 32252)
uwallface.dw <- get_download (32252)
# Get diatom counts for dataset 32252
uwallface.ct <- counts (uwallface.dw)
# Find the number of intervals and number of taxa in the count
sapply(uwallface.ct, dim)
# Look at the matrix of diatom counts
glimpse (uwallface.ct)
str(uwallface.dw) # str is a useful function to see the overall data structure for a dataset 
```

```{R}
#View sample metadata, list of taxa, and counts in Download dataframes 
#View sample metadata,
uwallface.dw[["32252"]][["sample.meta"]] 
#View taxa list
uwallface.dw[["32252"]][["taxon.list"]]
#View diatom counts
uwallface.dw[["32252"]][["counts"]]
#View Chronology
uwallface.dw[["32252"]][["chronologies"]]
```

```{r eval=FALSE}
library(neotoma)
library(neotoma2)
Names_744BH<-c("Bald Hill", "Blake (sutton)", "High (sudbry)", "Holland", "Little Hosmer", "Long (grnsbo)", "Long (shefld)", "Round (shefld)")
new_sites <- c()
for(name in Names_744BH){
sites <- get_sites(sitename=name)
new_sites <- c(new_sites, sites)}
new_sites

```
```{r}
library(neotoma2)
 Names_744BH<-c("Bald Hill", "Blake (sutton)", "High (sudbry)", "Holland", "Little Hosmer", "Long (grnsbo)", "Long (shefld)", "Round (shefld)")
 new_sites <- c()
 for(name in Names_744BH){
sites <- get_sites(sitename=name)
new_sites <- c(new_sites, sites)}
```

```{r}
library(neotoma2)
 Names_744BH<-c("Bald Hill", "Blake (sutton)", "High (sudbry)", "Holland", "Little Hosmer", "Long (grnsbo)", "Long (shefld)", "Round (shefld)")
 new_sites <- c()
 for(name in Names_744BH){
sites <- get_sites(sitename=name)
new_sites <- c(new_sites, sites)}
```

### List of diatom taxa - with name, authority, code, notes, etc.
```{R, eval=FALSE}
get_table(taxa)
get_taxa()
compile_taxa
```
Example of returning taxa 
```{R ListTaxaExample}
get_dataset(taxonids = )
get_dataset(datasettype = 'Diatom',taxonname = 'x')
```

### Water chemistry
```{R,eval=FALSE}
get_dataset(sitename='x',datasettype = 'diatom')
```
Example
```{R ChemistryExample}

```

### Chronology
```{R,eval=FALSE}
get_chroncontrol('x')# Can be thought like how bacon model is run through R
```
Example
```{R AnotherChronExample}

```

### Metadata - various options
```{R,eval=FALSE}
get_download()# Options within diatoms to organize it 
get_dataset()
```
Example
```{R DownloadStuffExample}
get_download()
get_dataset()
```

### Other types of data - LOI, geochemistry
```{R,eval=FALSE}
get_dataset(siteid='x')# Should be able to get the data from the siteid or dataset
```
Example
```{R}
get_dataset(siteid=,)
```


Issues:  Put metadata in the same files as the data matrices?  How to combine related files (e.g., diatom counts and chronology)? [When is it best to convert diatom counts to percentages?]

## Input files for data analysis programs
### Bacon
### Vegan
### Rioja
### Tidypaleo
### Others 

#Make plots and maps (e.g., stratigraphic diagrams, maps of taxa distributions)
###Plot of the abundance of multiple diatom taxa for a single core - vs depth or time
###Plot of abundance of a single taxon for cores from multiple sites - vs time
### Plot of abundance of a taxon vs an environmental characteristic (e.g., water chemistry variable)
### Map of distribution of sites that are part of an aggregate dataset
### Map of distribution of a diatom taxon; symbol size related to % abundance

#Modify and analyze data 
- and other miscellaneous tasks
## Ability to add to or modify data in Neotoma
```{R,eval=FALSE}

```
### add aggregate datasets to Neotoma
```{R,eval=FALSE}
set_dataset(x)# not code but having aggregate creation with set function sounds useful
```
Example
```{R}

```

### create new ecological groups and assign taxa to the groups
```{R,eval=FALSE}
set_taxa('x') # Not code but could be something
```
Example
```{R}

```

## Create and look for a table that explicitly matches water chemistry and diatom datasets.  This is especially important for cases where sites were sampled for water chemistry multiple times.
```{R,eval=FALSE}
# I think this could be an if and else statement since you have to compare 
```
Example
```{R}

```

## Extract a list of publications for a set of datasets
```{R,eval=FALSE}
<- get_publications(datasetid= x)
```
Example
```{R}


```

## Extract a list of DOI’s for a set of datasets
```{R,eval=FALSE}
get_publications(datasetid=)# Not code seems to need to be a mix of get_dataset and get_publications
```
Example
```{R}

```

## Create standard sets of diatom taxa names that could be used to convert multiple synonyms to a standard name
```{R,eval=FALSE}
get_dataset(taxonname = ,search())
```
Example
```{R}

```
