---
title: "Working with diatom data in the neotoma2 R package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with diatom data in the neotoma2 R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(neotoma2)
library(neotoma)
```

In the chunk to use neotoma::get_site(whatever) when using for Old neotoma package with new one 


One of the main goals of Neotoma is to pull records out of Neotoma into the R workflow.  

# Introduction
The ultimate purpose of this document is to provide general guidance and specific code examples for those wanting to use Neotoma R2 to retrieve and work with Neotoma diatom data.  An initial purpose is to help programmers (Simon, Socorro) develop Neotoma2 resources by providing a structured description of how diatom researchers may wish to use N-R2, and specific tasks they may wish to accomplish.

This document is organized to address two main goals of users, to browse the database to identify datasets that meet specified criteria, and to create output files that can be used for further analysis.  Other goals are to create plots and maps and to analyze data using R code.


#Browse data
The browse section describes ways of searching for available diatom datasets using a variety of criteria, either alone or in combination.  Criteria include those available in Neotoma Explorer, plus several more. Examples also include options for viewing and saving output results.

## Find diatom datasets using one or more search criteria. General task specification format: “Find all diatom datasets that meet the following criteria.”  [Chunks below show generic code specifying how to search / filter by a single criterion, followed by an example] 



###[Generic format of a complete chunk to browse using a single criterion]
```{R Diatom Generic Format, eval = FALSE}
diat_1 <- get_datasets(argument_1, argument_2)
```

#### Geographic region
```{R Diatom Geographic Region}

```
#####Geopolitical units (e.g., states)
```{R Diatom GPID}

```
#####Latitude and Longitude boxes
```{R Diatom Lat and Long, eval = FALSE}
diat_1 <- get_datasets(datasettype="diatom", loc=BOUNDING BOX)
```
#####Elevation
```{R Diatom Elevation}

```
#### Dataset type - Diatom, Diatom Surface Sample, Diatom Top Bottom
```{r getDiatoms, eval=FALSE}
neDiatoms <- get_datasets(datasettype="diatom surface samples", loc=BOUNDING BOX)
neDiatoms <- get_datasets(dois=c(...))
```



#### Site - Name(s) or ID(s)
```{r get_site,eval=FALSE}

```
#### Collection type
```{R Diatom Collection Type}

```
#### Constituent database
```{R Diatom Constituent Database}

```
#### Sample keyword
```{R Diatom Keywords}

```
#### Within or intersecting a specified time interval 
```{R Diatom Time Interval}

```
#### Member of one or more aggregate datasets
```{R Diatom Aggregate Dataset}

```
#### Related to one or more publications
```{R Diatom Publication}

```
#### From sites with specified range of water chemistry characteristics
```{r getWaterChem, eval=FALSE}
neDiatoms <- get_datasets(datasettype="diatom surface samples", loc=BOUNDING BOX)
```

#### From sites where one or more specified water chemistry characteristics were measured
```{R Diatom Multiple Water Chemistry}

```
#### Diatom Sample analyst name
```{R Diatom Sample Analyst}

```
#### Investigator name / person name
```{R Diatom Investigator}

```
#### Diatom counts include one or more specified diatom taxa
```{R Diatom Taxa}

```
#### Diatom counts include only taxa with abundance greater than x percent (in sample or core)
```{R Diatom Counts}

```

#### Site area larger or smaller than specified size
```{R Diatom Site Area}

```
#### Depository in which diatom slides are located
```{R Depository}

```
#### Chronology (y/n) and characteristics
```{R Chronology}

```

### Format for browsing using multiple criteria.  Examples: [could get complicated]  General rules and tips.  [Add as they become known]
#### Find all diatom datasets of any datatype
#### Find all diatom datasets of any datatype within a geographic region
#### Find all diatom surface sample datasets within a geographic region
#### Find datasets .... with surface area larger than x ha .......
### Other options / Misc:
#### Is water chemistry available?  (List variables included) Is chronology available?  Upper and lower age bounds.  Aggregate dataset(s) of which the dataset is a member.

## Create output of browse results.  Minimum “list” of dataset, collection, and / analysis unit ids for further browsing and use as input to other R functions. 
 
Standard comprehensive output file for use in detailed browsing. Variables should include all of the following (variables are grouped by their location on Tilia tabs):

### Site info:
#### Neotoma Site ID, Site name, Latitude, Longitude, Latitude width, Longitude width, Country, State, County, Admin Unit (3rd Geographic Unit), Lake characteristics

### Collection unit:
#### Collection unit type, Collection unit name, Collection unit handle, Location in site, Collection device, Date collected, Water depth of collection

### Dataset:
#### Dataset type - (Diatom, Diatom surface sample, or Diatom top-bottom), Primary publication, Investigators [or first author of publication], Sample analyst
```{R}

```
# Create data output files  
The create data output files section -  standard formats for diatom counts, metadata, chronology, water chemistry.  It provides examples for retrieving data and creating output files in a few basic formats.  Formats include those that can be used as input to data analysis programs - R packages and others.  In general, they should include all related data; the amount of data can be reduced later.They will include several descriptive variables and be easily readable. [input to code chunks can include output from browse tasks]

## Basic files / matrices
```{R}

```
### Diatom counts - single sites and multiple sites; aggregate data set; stratigraphic vs surface sample vs top-bottom; percentages vs raw counts.
```{R}

```
### List of diatom taxa - with name, authority, code, notes, etc.
```{R}

```
### Water chemistry
```{R}

```
### Chronology
```{R}

```
### Metadata - various options
```{R}

```
### Other types of data - LOI, geochemistry
```{R}

```

Issues:  Put metadata in the same files as the data matrices?  How to combine related files (e.g., diatom counts and chronology)? [When is it best to convert diatom counts to percentages?]

## Input files for data analysis programs
### Bacon
### Vegan
### Rioja
### Tidypaleo
### Others 

#Make plots and maps (e.g., stratigraphic diagrams, maps of taxa distributions)
###Plot of the abundance of multiple diatom taxa for a single core - vs depth or time
###Plot of abundance of a single taxon for cores from multiple sites - vs time
### Plot of abundance of a taxon vs an environmental characteristic (e.g., water chemistry variable)
### Map of distribution of sites that are part of an aggregate dataset
### Map of distribution of a diatom taxon; symbol size related to % abundance

#Modify and analyze data 
- and other miscellaneous tasks
## Ability to add to or modify data in Neotoma
```{R}

```
### add aggregate datasets to Neotoma
```{R}

```
### create new ecological groups and assign taxa to the groups
```{R}

```
## Create and look for a table that explicitly matches water chemistry and diatom datasets.  This is especially important for cases where sites were sampled for water chemistry multiple times.
```{R}

```
## Extract a list of publications for a set of datasets
```{R}

```
## Extract a list of DOI’s for a set of datasets
```{R}

```
## Create standard sets of diatom taxa names that could be used to convert multiple synonyms to a standard name
```{R}

```