---
title: "Working with diatom data in the neotoma2 R package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with diatom data in the neotoma2 R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(neotoma2)
library(neotoma)
```

In the chunk to use neotoma::get_site(whatever) when using for Old neotoma package with new one 


One of the main goals of Neotoma is to pull records out of Neotoma into the R workflow.  

### Diatoms

When working with diatoms there is surface and top-bottom and cores
* Pull diatom surface samples.
```{r getDiatoms, eval=FALSE}
neDiatoms <- get_datasets(datasettype="diatom surface samples", loc=BOUNDING BOX)

neDiatoms <- get_datasets(dois=c(...))
```


  * We may want to pull spatially, or pull by "National Assessment", or by other defined calibration set.

```{r getWaterChem, eval=FALSE}
neDiatoms <- get_datasets(datasettype="diatom surface samples", loc=BOUNDING BOX)
```


* Look for water chemistry at the same site (or collection unit).
```{r get__, eval=FALSE}

```
*Creating fill in place for header 

Look for a table that explicitly matches water chemistry and diatom datasets, in cases where records have been sampled multiple times.

* In theory collection dates are a good idea to use to filter, but collection dates might not be specific enough.

* In some regional studies water chemistry was sampled multiple times during the summer (e.g.) but diatoms were sampled only a single time.

* Was this going to be done in the aggregate dataset method?  Didn't happen.

Then:
* With one diatom record per water chemistry record
* we need taxon name, we need water chemistry type & units

Then we do some inference modeling (`rioja`).  Building e.g., transfer functions.

* Then run predictions on cores & plot predicted values & error

* We would extract an appendix of publications for the datasets we use.
* We would want some maps of site locations
* We would want some simple scatter plots
* We would want to do something about understanding taxonomic differences (so we'd want to know who the analyst was, or who the PI was)
* We might want to group or synonomize.  Can we pull those from Neotoma?
* We want the stratigraphic diagrams for cores, and surface sample diagrams for surface sample sets
* We want the lists of dataset DOIs.
* Want abundances of taxa along a gradient (with superimposed curves?)
