---
title: "Mammal Diversity in the Holocene"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mammal Diversity in the Holocene}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(neotoma2)
```

This is a model workflow to help understand and map assemblage diversity across time and space.

## Obtain Data
### Obtain all Neotoma Mammal Data

Download occurrence and age data from Neotoma: All vertebrate occurrences for XX region (North America, global) for a particular time window

```{r}
# This gives us 25 records (without setting 'limit')
mammals <- get_datasets(datasettype='vertebrate fauna')
```

### Obtain Secondary (Environmental) Data

Download climate data from appropriate source.

```{r}
worldTmin <- raster::getData('worldclim', var='tmin', res=10)
```

## Cleaning Data

### Clean ages

  * ensure taxa are really within the target age range
  * Assess what age controls are used for each species/occurrence?
  * Exclude certain dating methods (e.g. 14C on apatite, or biostratigraphic data, or… high grade based on the narrowness of the date control or error bars)

### Clean taxonomy

  *  Determine possible taxonomic level (species, genus)
account for uncertain taxa (cf., aff., type, ?) → taxa harmonization table

### Clean Spatial Records
  * For each species across all sites:
  * Check for geographic/environmental outliers

## Assemble Cleaned Data

  * Create convex hulls (at multiple time bins?)
Evaluate which convex hulls intersect with /come within a buffer of point of interest

### Join Hulls with Environmental Data
  
  * Apply the convex hull to look at the environmental space occupied by the taxon

## Build SDMs
  
  * Create SDMs for taxon with intersection convex hulls
  * Assess which SDMs intersect with point of interest
