---
title: "Mammal Diversity in the Holocene"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mammal Diversity in the Holocene}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup, eval=TRUE}
library(neotoma2)
library(raster)
library(dismo)
```

This is a model workflow to help understand and map faunal assemblage diversity across time and space.  There are more `neotoma2` vignettes, for diatoms and for pollen generally.  This vignette specifically addresses the use of the package to manage and understand how the package can be used to work with vertabrate faunal data.

## Obtain Data
### Obtain Neotoma Mammal Data

Download occurrence and age data from Neotoma: All vertebrate occurrences for XX region (North America, global) for a particular time window.

By default the API calls have a `limit` of 25 records at a time. Given this the call:

```{r getRecords, eval=TRUE, message=FALSE}
faunal <- get_datasets(datasettype = 'vertebrate fauna')
```

would return only the first 25 records in Neotoma.  We can extract the coordinates directly using the `coordinates()` function, we can plot the locations in a very simple `plot()` function, or we can create an interactive spatial plot using the `plotLeaflet()` command to display the records:

```{r plotSiteLatitudes, eval=TRUE}
faunal_coordinates <- neotoma2::coordinates(faunal)
plot(faunal_coordinates$lat, 
     ylab = "Latitude (degrees)", 
     xlab = "Longitude (degrees)")
```

The histogram we plot here allows us to examine the latitudinal gradient of the faunal records.  It includes only the 25 records we called for (the default) in the original `get_datasets()` call.  You can see how to call more records in the [Limits & Offsets section](#limits) below.

```{r plotSitesInteractive, eval=TRUE}
plotLeaflet(faunal)
```

The leaflet plot is useful because it provides an interactive element to the datasets (the map can be zoomed and clicked). The map is also a `leaflet` object, and so it can be further modified by the end user.  Note: The `leaflet` plot is not accessible to an individual who uses assistive technology to navigate the web.

### Limits & Offsets {#limits}

The default for all calls to the Neotoma API is to set a `limit` of 25, but often we want all records.  There are two ways of obtaining all the records.  One is to increase the `limit` parameter, for example, calling:

```{r getDatasetsHighLimit, eval=TRUE}
mam50 <- get_datasets(datasettype = 'vertebrate fauna', limit = 50)
```

One way of dealing with the issue of `offsets` and `limits` is the use of the `all_data` argument. When set to `TRUE`, a `while` loop is run within the function, to pull the data responsibly. This will result in returning all the data for your use-case.

```{r downloadMammals, eval=FALSE}
# This gives us 25 records (without setting 'limit')
mammals <- get_datasets(datasettype = 'vertebrate fauna', limit = 200)

# The 'all_data' approach:
allMammals <- get_datasets(datasettype = 'vertebrate fauna', 
                           all_data = TRUE)
```

In general it is better to filter records using the search parameters available within the `get_*()` functions than to download all records and filter within R.

For any query set returned by the API we can examine the objects we have returned:

```{r summaryExample}
summary(allMammals)
```

For our example, we want to search for all early Holocene vertebrate records within Neotoma in eastern North America.  So we can construct our search as follows:

```{r getEasternFossils, eval=FALSE}
mamSet <- get_datasets(datasettype = "vertebrate fauna",
                       ageyoung = 6000,
                       ageold = 11500,
                       loc = c(-100,20,20,90),
                       all_data = TRUE)
```

The `get_datasets()` command does not explicitly download the raw data.  `get_datasets()` is intended to provide an outline of the data available.  It is then used with the `filter()` functions (described in [Cleaning Data](#cleaning)), and ultimately with `get_downloads()` to return the raw data.

## Cleaning Data {#cleaning}

The basis of any workflow is downloading appropriate data. Lets begin again by downloading all records for mammals within the database that have a late-Pleistocene age.  Here we use the `get_datasets()` command, and set `ageold` to be `12000`.  Dates passed to the API are assumed to be in calibrated radiocarbon years before present.

```{r obtainMammalData, eval=TRUE}
mamSet <- get_datasets(datasettype = "vertebrate fauna",
                       ageyoung = 6000,
                       ageold = 11500,
                       loc = c(-100,20,20,90),
                       all_data)

plotLeaflet(mamSet)
```

With the leaflet plot we can dynamically navigate records, and check to see if sites meet our criteria by checking the records against the Neotoma dataset landing pages, or Neotoma Explorer.

We can use the `filter()` function to remove data based on one or more criteria.  For example, if we think a particular site is not appropriate to our analysis we can filter by `sitename` before downloading the records:

```{r filterandDl}
mamDl <- mamSet %>% 
  neotoma2::filter(sitename != 'Medicine Creek Site [39LM222]') %>% 
  get_downloads(., all_data = TRUE)
```

The full set of filter arguments supported are:

 | -------- | ------------ | ----------- |
 | "siteid" | "collunitid" | "datasetid" |
 | "sitename" | "lat" | "long" |
 | "area"  | "sitenotes" | "description" |
 | "altitude" | "database" | "datasettype" |
 | "age_range_old" | "age_range_young" | "datasetnotes" |
 | "handle" | "colldate" | "location" |
 | "waterdepth" | "collunittype" | "collectiondevice" |
 | "defaultchronology" | "collectionunitname" | "depositionalenvironment" |

Once the data is filtered we can either continue to manage the data as sites, datasets or collection units, or we can begin to work directly with the raw sample data using the function `samples()`.  The `samples()` function 

### Clean ages

Given a set of downloaded datasets we can make additional choices and decisions about the records.  We know that in Neotoma there are a number of records that are measured using radiocarbon years, while others use calendar years.  We also know that datasets have temporal bounds that may extend beyond the ranges of the samples we're interested in (for example, a site has samples across the entire Holocene, but we're only interested in a narrow time window around a certain event).

So, given the 

```{r getSamples}
mamSamp <- samples(mamDl)
```

  * ensure taxa are really within the target age range
  * Assess what age controls are used for each species/occurrence?
  * Exclude certain dating methods (e.g. $^{14}C$ on apatite, or biostratigraphic data, or… high grade based on the narrowness of the date control or error bars)

### Clean taxonomy

  *  Determine possible taxonomic level (species, genus)
account for uncertain taxa (cf., aff., type, ?) → taxa harmonization table

### Clean Spatial Records

  * For each species across all sites:
  * Check for geographic/environmental outliers

## Assemble Cleaned Data

  * Create convex hulls (at multiple time bins?)
Evaluate which convex hulls intersect with /come within a buffer of point of interest


### Obtain Secondary (Environmental) Data

Download climate data from appropriate source.

```{r, eval=TRUE}
worldTmin <- raster::getData('worldclim', var = 'tmin', res = 10)

raster::plot(worldTmin, 1)
```

This is just an example of downloading one element of [WorldClim](https://www.worldclim.org/) data as a raster in R. This particular raster has `r raster::nlayers(worldTmin)` layers and is a global dataset.  The `raster` package for R has a number of links to external datasets such as WorldClim.  Alternately you can load in your own data.

### Join Hulls with Environmental Data
  
  * Apply the convex hull to look at the environmental space occupied by the taxon

## Build SDMs
  
  * Create SDMs for taxon with intersection convex hulls
  * Assess which SDMs intersect with point of interest
