---
title: "The neotoma2 R Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The neotoma2 R Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(neotoma2)
```



## Publications

Many Neotoma records have publications associated with them.  The `publication` object (and the `publications` collection) provide the opportunity to do this.  The [`publication`](http://open.neotomadb.org/dbschema/tables/publications.html) table in Neotoma contains an extensive number of fields.  The methods for `publications` in the `neotoma2` package provide us with tools to retrieve publication data from Neotoma, to set and manipulate publication data locally, and to retrieve publication data from external sources (e.g., using a DOI).

### `get_publications()` from Neotoma

Publications in Neotoma can be discovered using a number of parameters.  
