---
title: "Neotoma2 R-package - JASM Workshop"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The neotoma2 R Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(neotoma2)
library(dplyr)
library(purrr)
library(sf)
library(tidyverse)
```


## Outline
### Browse Data
#### Find sites
#### Find datasets
### Download data
#### Download full datasets
#### Subset download output
#### Create output files
### Visualize data
#### Tables [?]
#### Maps
#### Diagrams



## Introduction
This document helps introduce workshop participants to use of the neotoma2 R package, primarily to browse, download, and analyze data in the Neotoma database.  It provides general guidance and specific code examples for those wanting to use `neotoma2` to retrieve and work with data (primarily diatom-related) from the Neotoma Paleoecology Database. 

A main goal of `neotoma2` is to pull records from the Neotoma database and into a user's R workflow. This will help diatomists and other environmental scientists produce clear working documents that use data from Neotoma and allow them to combine it with their own data or data from other sources.

This document is organized to address users' two main goals: to browse the database to discover datasets that meet specified criteria and to create output files that can be used for further analysis. Other goals are to create plots and maps and analyze data using R code.

### Data Organization in Neotoma

It is worth knowing how the Neotoma database is organized and what we mean when talking about things like `sites`, `datasets`, and `samples`.

The Neotoma data model is available online at the [Database Schema](https://open.neotomadb.org/dbschema) page. As of September 2021, there were 150 different tables in the database. Some of these have particular kinds of data (for example, `agetypes`), and some link tables together. The `neotoma2` package directly supports access to some tables, but not all. For specialized cases, a user may be required to access an entire table directly (using the `get_table()` function).

#### Sites, Datasets, etcetera

Neotoma data is structured spatially. Each collection location is assigned a **site**, a spatial location defined by either a point or a polygon. It may be a lake, an archaeological site, a swamp, or another location.

Within a **site**, there may be one or more **collection units**.  
A collection unit represents the material (e.g., sediment core) removed from a specific location within a `site` (e.g., a lake). So, a researcher may have taken multiple cores within a lake basin, or an archaeologist may have several dig sites within a larger area. Collection unit material is subdivided into **analysis units** (e.g., sediment intervals within a core). **Samples** are defined as material removed from analysis units and analyzed for a particular dataset type (e.g., diatoms, pollen, water chemistry). Groups of analysis units (samples) from an individual collection unit are then grouped into a **dataset**.

#### What is the R package doing?

The `neotoma2` package uses data from the Neotoma database, but it is mediated through R functions (like `get_sites()` to get site information) that call public URLs, which are abstractions of database calls. So, for example, to get data for the site with ID number 12:
```r
get_sites(12)
```
calls out to the internet for the result at:
```
https://api.neotomadb.org/v2.0/data/sites/12
```

Which itself calls an SQL query:
```sql
SELECT * FROM ndb.sites
WHERE siteid = 12
```

It is a lot easier for most users to use the R functions since they create objects in R that work with other commands and statistical analysis they might want to do.
 

##### `neotoma2` functions

There are several functions in `neotoma2` to retrieve and subset data.  Functions used in this workshop document include `get_sites()`, `get_datasets()`, `get_downloads()`, filter(), collunits(), datasets(), samples(), and plotLeaflet().  See the neotoma2 package documentation for more details on these and other functions.[[These are described where they first appear in the document below.]]


## Browsing - Data discovery

This section describes ways of searching the Neotoma database to find `sites` and diatom `datasets` using various criteria. The primary purpose of each code chunk is to find the `sites` or `datasets` that meet the criteria specified in the code. The returned objects can then be used as input to other code chunks that return more detailed information on the datasets and create output files for data analysis (e.g., diatom counts).

### Find sites
Use the get_sites or get_datasets functions to find sites and related information.  Both functions return essentially the same information, but searches using siteid or sitename can only be done using get_sites.
Both functions can be used to search for information on a single or multiple sites. The functions are helpful for finding locations of all sites within a region, even if they are not diatom sites; this is useful for indicating the level of paleoecological 'effort' in a region.  Also, output can be used to generate a map of all sites and superimpose specific sites of interest.

#### get_sites function
**Usage:** Sites can be searched using the following supported parameters.
e.g., `get_sites(sitename="Alexander Lake")`

| parameter | allowed values |
| ------ | ------- |
| `siteid` | integers (values > 0) |
| `sitename` | valid character string; `%` as wildcard |
| `altmax` | integers (values >= than `altmin`) |
| `altmin` | integers |
| `loc` | spatial coordinates (WKT or geoJSON) |
| `limit` | integer > 0; default 25 |
| `offset` | integer > 0; default 0 |

`limit` indicates the maximum number of records that will be returned; `offset` indicates the number in a sequence where processing will begin.  Together, these allow large numbers of records to be processed in batches.

**Values returned:** A "sites object", which includes nested lists with information on: siteid, sitename, latitude and longitude, elevation, geopolitical unit, notes, description, collection units, and datasets associated with the collection units.  In addition to the data above, the function returns a table (dataframe) with 5 location variables for each site.

To see the structure (data model) of `get_sites()` output, run the following code chunk.

##### Search by site ID / view structure of get_sites output

Search for a site by siteid and view the structure of the output.  
```{R site-id_structure_of_get_site_output}
uwallf_site <- get_sites(745) # finds site data for Upper Wallface Pond, which has Neotoma siteid 745.  )
str(uwallf_site) # shows structure of the uwallf_site object
```

##### Search by site name

If we know the names of one or more of the existing sites we are looking for, we can get site data using the `sitename` parameter.

Get data on a single site using sitename.
```{r sitesbyName example}
# diat <- get_sites(sitename = "")  # Basic format
diat <- get_sites(sitename = 'Schrader Pond')  # Example
diat
# wildcards (%) can be used if you do not know the exact name
## Example: get_sites(sitename = '%Schra%')
```

Get data on multiple sites using sitenames.

```{r multsitesbySitename example}
# Join a list of site names together using the combine command, c()
sitenameList <- c("Bald Hill", "Blake (sutton)", 
                  "High (sudbry)", "Holland", 
                  "Little Hosmer", "Long (grnsbo)", 
                  "Long (shefld)", "Round (shefld)")
my_sites <- c()
for(sitename in sitenameList){
  a_site <- get_sites(sitename = sitename)
  my_sites <- c(my_sites, a_site)
}
my_sites
```
##### Search for sites in a latitude and longitude bounding box

Example of get_sites by location with bounding box of lat & longs.  Same code works with get_datasets as well as get_sites.

```{R diatomsbyBoundingBox}
# Find sites in the northeast US.  Order of boundaries is: West, South, East, North.  Data are returned for the first 25 sites.
diat_NE <- get_sites(loc =c(-80,40,-65,48)) 
```

##### Find datasets
###### get_datasets function
**Usage:** get_datasets(x, datasettype, piid, altmin, altmax, loc, gpid, taxonids,taxonname, ageold, ageyoung, ageof, subdate)

x is an optional value that represents an object with numeric dataset ID(s). It can be a single datasetid, an array with multiple IDs, or an object created using get_sites.

**Values returned**`A "sites object" plus a dataframe with location information the same as that produced by the get_sites function.[[but with a more complex detail at the dataset level.  [[[Confirm this description!!  How is output different??]]]

To see structure of a`get_datasets()` object, run the following code:
```{R structure_of_get_dataset_output}
uwallf_site <-get_sites(sitename = 'Upper Wallface%') # finds site data for a site
uwallf_datasets <- get_datasets(uwallf_site) # finds datasets data for Upper Wallface (returns 4 elements corresponding to the 4 datasets available for U Wallface Pond:  diatom surface sample, diatom, chronology and water chemistry)
str(uwallf_datasets)
```

Don says: add table showing siteid, sitename, collection unit id, datasetid, datasettype, age_range_old, age_range_ young, first publication
----------------
```{R structure_of_get_dataset_output}
#SITE DATASETS
SITEID <- rep(uwallf_dl[[1]]@siteid,(length(as.data.frame(datasets(uwallf_data))$datasetid)))
SITENAME <- rep(uwallf_dl[[1]]@sitename,(length(as.data.frame(datasets(uwallf_data))$datasetid)))
DATASETID <- as.data.frame(datasets(uwallf_data))$datasetid
#DATASETNAME <-
DATASETTYPE <- as.data.frame(datasets(uwallf_data))$datasettype
AGE_RANGE_OLD <- as.data.frame(datasets(uwallf_data))$age_range_old
AGE_RANGE_YOUNG <- as.data.frame(datasets(uwallf_data))$age_range_young
#FIRST PUBLICATION <- (have to figure out) get_publications(DATASETID, 

data.frame(SITEID,SITENAME,DATASETID,AGE_RANGE_OLD,AGE_RANGE_YOUNG)
```
-------------------

#### Dataset ID
Get data on a single dataset using datasetID 

```{r get datasetID}
# diat_ID <- get_dataset(datasetid) - basic format
diat_ID <- get_datasets(44764) # Example
diat_ID
```

Get data on multiple datasets using multiple dataset IDs 
```{r multdatsetid }
multid <- c(44764,44765,44766)
diat_multid <- get_datasets(multid) # Example
diat_multid
```

#### Elevation (m)
Find datasets for sites between minimum and maximum elevations

```{R getDatasetssByElevExample}
diat_el <- get_datasets(altmin=500,altmax=5500)
diat_el
```

###### get_downloads
Using the dataset ID, site object or dataset object, return all records associated with the data as a `sites` object with an even more complex detail level.

**Usage:**  `get_downloads(x)`
Arguments: `x` can be a dataset ID or a vector of numeric dataset IDs. It can also be a `sites` object as returned by `get_datasets()` or `get_sites()`; verbose, logical printing a download status bar.

The values are returned as a `sites` object with a full level of detail; relevant assemblage information and metadata.

The downloaded result is a sites object that contains the constituent taxa, the chronology, site and PIs who contributed the data. 

The following are important components: datasets, chronologies, samples within the dataset.

To see structure of `get_downloads()` function output, run the following code:
```{R structure_of_get_download_output}
uwallf_site <-get_sites(sitename = 'Upper Wallface%') # finds site data for a site
uwallf_datasets <- get_datasets(uwallf_site) # finds datasets data for Upper Wallface (there's 4 elements)
uwallf_diatom_dataset <- neotoma2::filter(uwallf_datasets, datasettype == "diatom") # only keep diatom datasets
uwallf_download <- get_downloads(uwallf_diatom_dataset)
#str (uwallf_download) # shows structure of uwallf_datast
```


Get Sample IDs for Dataset IDs

Insert table listed on table doc (Dataset diatom counts / water chemistry)

## Using neotoma2::filter

There might be cases when we need to filter results out from a `sites` object. We can use the `filter()` function. `filter()` takes any `sites` attribute to filter on. lat, long, elev, datasettype

Don says: This should be moved up on the document (where this is first used)
-----
From Neotoma2.Rmd:
The `filter()` function takes as the first argument, a datasets object, followed by the criteria we want to use to filter. Current supported criteria includes:

* `lat`
* `long`
* `elev`
* `datasettype`

You also need to make sure that you accompany any of these terms with the following boolean operators: `<`, `>` or `==`, `!=`. Datasettype has to be of type string, while the other terms must be numeric.  If you need to filter by the same argument, let's say, you need to filter "geochronologic" and "pollen datatypes, then you will also make use of `&` and `|` operators.
----
#### Within or intersecting a specified time interval 
[Simon - Code still needs to be written]
Don says: Find a working example, or delete this section. Should be using age_old and age_young.

```{R Diatom Time Interval}
datasets(NL2)
NL2
```

```{R neotoma version}
interval_42 <- NL2 %>%
  neotoma2::filter(age_range_old > -42)
interval_42
```

```{r downloads after dplyr}
get_downloads(interval_42)
```


#### Taxon name or ID 
** Note taxon name / id not sure if in API)
```{R getTaxa}
# Find datasets using taxon name
diat_nm<- get_datasets(taxonname = 'Navicula minima')
# Find datasets using taxon id
```

```{r datasettype and filter taxons}
# Find datasets of a specific data type
diat_nm1<- get_datasets(datasettype = 'diatom')

# Get the downloads detail to retrieve the taxons. Apply samples() to the downloads object
diat_nm1_dw <- get_downloads(diat_nm1)
diat_samples <- samples(diat_nm1_dw)
diat_samples

# Use dplyr::filter to obtain the taxa of interest
diat_samples <- diat_samples %>%
  dplyr::filter(variablename == 'Navicula minima')
```

#### Dataset DOI
Don says: keep in, but make code work.
** Not sure dois is a used parameter. 
```{r getDiatomsByDOI}
# neDiatoms <- get_datasets(dois=c(...)) # basic format
neDiatoms <- get_datasets(dois=c(43516, 43519, 46228)) # Example using DOI's for Livingston P, West P and Wolf L in the Adirondack Mountains, NY.
# [SIMON - This code chunk does not work.  I assume this is because the Neotoma2 get_datasets function is still being developed. True?  My understanding is that it would retreive the DOI. Don]
```

### Find datasets (multiple functions)
#### Constituent database
Don says: keep searching by keyword, but get rid of constituent database example (?)
[SIMON - Code needs to be added]
```{R SitesbyConDB}
consDB <- get_
```

Example of finding constituent databases

```{R SitesbyConDBExample}
ConDB <- 
  
  ```

#### Sample keyword [secondary]
Code to retreive datasets by keyword - usually would be one part of a larger code chunk or function
[Code needs to be written]

Example of using the search function returning keywords 
```{R KW in notes}
sediment_txt <- as.data.frame(datasets(diat_nm1_dw)) %>%
  dplyr::filter(grepl('Sediment sample', notes))

sediment_txt$datasetid

```

#### Aggregate dataset(s) <- ??
Don says: Yes, we want this section.
Get dataset IDs of all members of an aggregate dataset
Also, get_datasets and get-downloads
```{R Diatom Aggregate Dataset}
aggr_set <- read.csv()# If we are going to use the Excel file that contains sampleids for each aggregate dataset, we should read it in here
aggr1 <- get_datasets() # Do we need to use get_datasets if we already know sampleids?
aggr2 <- get_downloads()
# Code needs to be written
```

Example
```{R Diatom Aggregate Dataset Example}
```

#### Water chemistry
Don says: We want water chemistry
##### Is water chemistry available?
Determine if any water chemistry data is available for one or more sites or diatom datasets for those sites
If chem data are available, which parameters? (List variables included)
[SIMON - Code chunks in this section do not work]

##### From sites within range of water chemistry characteristics
Get datasets associated with water chemistry within a specific range (e.g., pH between 6 and 8)

Example of how to get water chemistry from the different dataset types [includes snippets of code, none of which work]

```{r getWaterChemEX}
diatomss <- get_datasets(datasettype = 'diatom surface sample') %>% 
  get_downloads()

# Following 3 lines of code seem to work, but search through ALL water chemistry datasets
waterchem <- get_datasets(datasettype = 'water chemistry') %>% 
  get_downloads() 
```

**Note Socorro: not clear what is supposed to happen here; dont find any property called pH so tried with waterdepth.
```{r}
as.data.frame(collunits(waterchem)) %>% 
  dplyr::filter(waterdepth < 5)

# Below - for the future....  get_aggregate function does not yet exist
diatomss <- get_aggregate(123)
waterss <- get_aggregate(124)
diatomdl <- get_downloads(diatomss) %>% compile_downloads()
waterdl <- get_downloads(waterss) %>% compile_downloads()
# Following proposed function would match corresponding diatom and water chemistry samples
aligned <- align_sites(diatomss, waterss)
```

## Need explanations on here (?)
```{r fixing}
diatomss<- get_dataset(datasettype = 'diatom surface sample')
waterss<- get_dataset(datasettype = 'water chemistry')
diatom<- get_dataset(datasettype = 'diatom')
#Sapplyy function is used to return diatomss, read / extract the metadata and select dataset IDs for the first 20 lakes 
subsetss <- sapply(diatomss, 
                   function(x) {
                     x$dataset.meta$dataset.id
                   })[1:20]
#Sapplyy function is used to return diatom, read / extract the metadata and select dataset IDs for the first 20 lakes
# [Simon - I am not sure what the following code does - please document] ?
subset <- sapply(diatom, 
                 function(x) {
                   x$dataset.meta$dataset.id
                 })[1:20]
if('filename' %in% list.files()) {
  diat_dl <- readRDS('filename') # It returns the file that is used 
} else {
  diat_dl <- get_download(subset)
  diat_dlss <- get_download(subsetss)
  diat_dl <- bind(diat_dl, diat_dlss)
  saveRDS(diat_dl, 'filename') # Saves the download object as a file
  #  Where is "filename"saved? - What directory? - type getwd() in the console to find out where the file is.
}
```



##### From sites where one or more specified water chemistry characteristics were measured
```{R Diatom Multiple Water Chemistry}
library('WriteXLS')
diatomwch<- get_datasets(datasettype = 'water chemistry')
diatom<- get_datasets(datasettype = 'diatom surface sample')

subsetwch <- getids(diatomwch)$datasetid[1:20]

subset <- getids(diatom)$datasetid[1:20]

if('fileWCH' %in% list.files()) {
  diat_dl <- readRDS('fileWCH') # It returns the file that is used 
} else {
  diat_wch <- get_downloads(subsetwch)
  diat_dls <- get_downloads(subset)
  diat_dlb <- c(diat_wch, diat_dls)
  saveRDS(diat_dlb, 'fileWCH')
}
# fileWCH will be created in the working directory
#wideWCH_DTM <- compile_downloads(diat_dlb)  # what is this supposed to do?
## neotoma function converts multiple downloads into single object 
minWCH <- as.data.frame(collunits(diat_dlb))

m_ids <- getids(diat_dlb)

minWCH <- left_join(minWCH, m_ids, by = c("collectionunitid" = "collunitid")) %>%
  group_by(datasetid) %>% #dataset was returned from the table 
  summarise(min(waterdepth)) %>% #dplyr function to create a dataframe
  rename(waterdepth = "min(waterdepth)")

wideWCH_DTM <- minWCH %>% 
  filter(is.na(waterdepth))
wideWCH_DTM
# Write data to an Excel file, which will be created in the working directory
WriteXLS(wideWCH_DTM, "wideWCH_DTM.xlsx")
```
## Socorro from here April 4th

#### Diatom taxa
Don says: We probably want this.
##### Diatom counts that include one or more specific diatom taxa
[SIMON - Code does not include possibility of specifying a particular diatom taxon]

Get list of taxa found in a single dataset (core)

```{R datasetByTaxa}
diat_tax <- get_datasets(32252) # Datasetid for U Wallface P core 1
diat_dl <- get_downloads(diat_tax)
diat_d2 <- compile_downloads(diat_dl)
diat_taxa <- taxa(diat_dl)
diat_taxa
```

Get list of all taxa associated with single datasettype 

```{R datasetByTaxaExample}
diatoms<- get_dataset(datasettype = 'diatom')
diat_dl <- get_download(diatoms)
diat_taxa <- taxa(diat_dl) # List of taxa in all datasets of dataset type "diatom"
# Takes a long time to run
```


##### Diatom counts include only taxa with abundance greater than x percent (in sample or core)

```{R Diatom Counts}
```

Example

```{R Diatom Counts EX}
```


### Format for browsing using multiple criteria - Examples 
Examples: [could get complicated]  General rules and tips.  [Add as they become known]
### [Following sections to be reorganized]
#### Find all diatom datasets of any datatype

#### Find all diatom datasets of any datatype within a geographic region

#### Find all diatom surface sample datasets within a geographic region

#### Find datasets .... with surface area larger than x ha .......

### Other options / Misc:

### Create output of browse results 
#Don says: old version of data table doc, leave for now
Minimum “list” of dataset, collection, and / analysis unit ids for further browsing and use as input to other R functions. 

Standard comprehensive output file for use in detailed browsing. Variables should include all of the following (variables are grouped by their location in Tilia tabs): 
1.  Site information: Site ID, Site name, Latitude, Longitude, Latitude width, Longitude width, Country, State, County, Admin Unit (3rd Geographic Unit), Lake characteristics, publication id
2. Collection unit information:  CU ID, Type, Collection unit name, Collection unit handle, Location in site, Collection device, Date collected, Water depth of collection
3.  Dataset information: Data Type (Diatom, Diatom surface sample, or Diatom top-bottom), Primary publication, Investigators [or first author of publication], Sample analyst


## Create output files for data analysis

Code chunks in this section are designed to create output files for diatom counts, metadata, chronology, water chemistry, and other kinds of data.  They provide examples for retrieving data and creating output files in a few basic formats.  Formats include those that can be used as input to data analysis programs - R packages and others.  In general, though they do not do so now, basic output files should include all related data; the number of variables can be reduced later.They will include several descriptive variables and be easily readable. [input to code chunks can include output from browse tasks]

### Basic files / matrices

#### Diatom counts
The following code chunks can be used to view diatom count related data.  There are a few examples, each with something a little different.
For single sites and multiple sites; aggregate data sets; stratigraphic vs surface sample vs top-bottom; percentages vs raw counts; containing specific taxa, etc.

#Stopped here: April 8, 2022 - 3:51 pm
##### View sample and count metadata
```{R Upper Wallface View counts}
### View sample metadata, list of taxa, and counts in Download dataframes 
### View sample metadata,
uwallface.dw[["32252"]][["sample.meta"]] 
### View taxa list
uwallface.dw[["32252"]][["taxon.list"]]
### View diatom counts
uwallface.dw[["32252"]][["counts"]]
### View Chronology
uwallface.dw[["32252"]][["chronologies"]]
```
##### Count data for one or more datasetids
Gets count data for one or more datasets and writes data to an Excel file.
```{R ExampleDataset}
library(purrr)
V1 <-c(44786,44787,44788) # Datasetids
# In this example we combine datasets from 3 sites (New England lakes).  To get output for aggregate datasets, a file with all the datasetids for an aggregate dataset could be used as input.
V1_diatom <- map(V1, function(x)
  neotoma:::get_dataset(x,datasettype="diatom")) %>%
  neotoma::bind()   
#In the get_dataset line, whatever is in the parenthesis (a parameter such as x or gpid); will be whatever is in the object V1, defined in the line above.
# [SIMON - Why does datasettype need to be specified above, given that input data are datasetids?  The code seems to run fine with "datasettype="diatom" removed.]
if('V1' %in% list.files()) {
  V1_dl <- readRDS('V1') 
} else {
  V1Datasets <- list()
  #
  for(i in 1:length(V1_diatom)) {
    V1Datasets[[i]] <- get_download(V1_diatom[[i]])
  }
  #For loop will go through each step to first create the file and then download data for each site
  V1_download <- neotoma::bind(V1Datasets)
  saveRDS(V1_download, 'V1')
}
V1Counts <- compile_downloads(V1_dl)
V1Row <- V1Counts %>% 
  group_by(dataset) %>% 
  summarise(min(depth))
WriteXLS(V1Counts,  "VL_dl.xlsx")
#This last step creates a table in an Excel file which will be saved in the working directory
```

##### Count data file for a single site using site name
Following code uses Neotoma2 R - Schrader Pond example
[SIMON - code does not run - "Error in get_downloads(schrader.data):could not find function "get_downloads"".  I assume this is because the Neotoma2 get_downloads function has not yet been made available]
```{r get_singleSitebyName}
schrader <- get_sites(sitename = 'Schrader%') # % can be used to approximate sitenames
schrader  # Look in output to find datasetid for Schrader (27408)
schrader.data <- get_datasets(27408)
schrader.dl <- get_downloads(schrader.data)
schrader.cnt <- counts(schrader.dl)
sapply(schrader.cnt, dim) 
print(schrader.dl)
```

#### Count data for one dataset ID - subsets

## [SIMON - please document the code below - what is its purpose?]
if('Mirror' %in% list.files()) {
  diat_da <- readRDS('Mirror') 
} else {
  diat_dla <- get_download(subsetda)
  saveRDS(diat_dla, 'Mirror')
}
#The if and else loops work together to create a file containing downloaded data for the subset
diatCount <- compile_downloads(diat_dla) 
## SIMON - please explain the purpose of the following code.  Seems like maybe this overall code chunk was designed to be able to do more than just get count data for one surface sample.
minimumRow <- diatCount %>% 
  group_by(dataset) %>% 
  summarise(min(depth)) 
#In compiling downloads, information is displayed in a table format and is organized using the group_by and summarise functions
```

##### Count data for a single site name or datasetid
MAYBE revise to neotoma2 to use as the basis for a workshop example.

Upper Wallface Pond examples
```{R Upper Wallface EX}
library (neotoma)
### Find Neotoma Site Id for Upper Wallface Pond 
uwallface <- get_site (sitename = "Upper Wallface%")
### Find datasets and IDs associated with U Wallface Pond
uwallface.dl<- get_dataset(uwallface)  # Review datasets available note ID of the one of interest
### download data for diatom core dataset (Dataset 32252)
uwallface.dw <- get_download (32252)
### Get diatom counts for dataset 32252
uwallface.ct <- counts (uwallface.dw)
### Find the number of intervals and number of taxa in a count
sapply(uwallface.ct, dim)
### Look at the matrix of diatom counts
glimpse (uwallface.ct)
uwallface_comp <- compile_downloads(uwallface.dw) 
# Compile_downloads is used for combining multiple counts into one table
uwallface_comp
WriteXLS(uwallface_comp,"UWallDtmCount.xlsx", col.names=TRUE, row.names=TRUE)
```


#### List of diatom taxa  [[working code still needs to be written - dfc]]
Don says:  replace this code chunk with code using the "taxa" function in Neotoma 2.
[Simon - following code does not work - (Error: object 'x' not found)
Create file of all diatom taxa in the Neotoma master list - with name, authority, code, notes, etc.
```{R, diatom taxa out}
x
Taxa_hi <- sapply(x, 
                  function(x) {
                    x$taxa
                  })
compile_taxa
```

```
Get count data for a single core
```{R Counts single dataset UWall EX}
library (neotoma)
### Find Neotoma Site Id for Upper Wallface Pond 
uwallface <- get_site (sitename = "Upper Wallface%")
### Find datasets and IDs associated with U Wallface Pond
uwallface.dl<- get_dataset(uwallface)  # Review datasets available; note ID of the one of interest
### download data for diatom core dataset (Dataset 32252)
uwallface.dw <- get_download (32252)
### Get diatom counts for dataset 32252
uwallface.ct <- counts (uwallface.dw)
### Find the number of intervals and number of taxa in a count
sapply(uwallface.ct, dim)
### Look at the matrix of diatom counts
glimpse (uwallface.ct)
str(uwallface.dw) # str is a useful function to see the overall data structure for a dataset 
Comp_UWall <- compile_downloads(uwallface.dw) # used for formatting into a table 
WriteXLS(uwallface.ct,  "count_file.xlsx")
```
#### Site information and environmental data
Needs to be written

#### Water chemistry
```{R WCHM Out}
## get_datasets(sitename='x',datasettype = 'water chemistry') # format
get_datasets(sitename="%Wallface%",datasettype = 'water chemistry') # example
```
Example
```{R ChemistryExample}
```
#### Chronology
Don says: replace with code using "chronologies" function in Neotoma2.
One way to get chronology data is to first run get_download to create a download file and then run get_chroncontrol to extract the chronology data.  The example below gets chronology data from  download files with data for all sites in the northeast US.
```{R ExtractChroncontrol}
# Get chroncontrol data from a download file using get_chroncontrol
ne_chroncontrol <- get_chroncontrol(ne_down_dtsts, chronology = 1, verbose = TRUE, add = FALSE)
```
```{R ChronMetadata}
# Get chronology metadata from the .meta and .parent lists
ne_chronmeta <-sapply (ne_chroncontrol, 
                       function(x) {               c(x$meta$name, x$meta$chron.id, x$meta$age.type, x$parent$dataset.name, x$parent$dataset.id, x$parent$dataset.type)
                       })
str(ne_chronmeta)
```

### Input files for data analysis programs
Some programs have no standard input format
#### Bacon, ### Vegan, ### Rioja, ### Tidypaleo
#### Others 



### Add aggregate datasets to Neotoma

### Transform raw diatom counts to percentages and other forms
Use the tran function in the Analogue package to transform raw data.  See analogue package documentation for a list o possible transformations.

``` {R TransformCountsToPct}
library ("analogue")
uwallf_dtm_pct <- analogue::tran (x = uwallf_download [[1]]$counts, method = "percent") # Transforms raw counts to percents; uwallf_download is a get_ download filew with diatom counts for Upper Wallface Pond
```

Note: Basic code for above chunk is from S. Goring, J. Williams and E. Grimm Neotoma R Workshop documentation

### Add to a table that explicitly matches water chemistry and diatom datasets
This is especially important for cases where sites were sampled for water chemistry as part of more than one study.  As of Oct 2021, corresponding water chemistry and diatom counts are stored in an Excel file.

### Extract a list of publications for a set of datasets
Returns publicationid, citation, and DOI for pub if there is one
```{R get pub out}
# get_publications(datasetid= x)  # Basic format
uwall_pubs <- get_publications(datasetid=32252) # Example - U Wallface P datasetid
uwall_pubs
```

### Extract a list of DOI’s for a set of datasets
```{R DOIS}
# Code needs to be written
```
Example
```{R DOIS EX}
```

### Add standard sets of diatom taxa names that could be used to convert multiple synonyms to another set of names
A complicated task - synonym lists need to be prepared and code written
```{R standard set}
get_dataset(taxonname = ,search())
```

## Make plots and maps 
(e.g., stratigraphic diagrams, maps of taxa distributions)

### Map of distribution of sites 
Map, for example, all sites that are part of an aggregate dataset. Following example uses Leaflet package. Very useful for making basic maps quickly and easily.

-----------------------------
-----------------------------
**ANY material below these lines will not be included in the final workshop document**

### Map of distribution of a diatom taxon
Option 1 - basic map showing occurrence of a taxon
Option 2 - basic map showing occurrence of multiple taxa
Option 3 - same as option 1, but with symbol size proportional to % abundance
Option 4 - same as option 2, but with symbol size proportional to % abundance

### Plot of the abundance of diatom taxa in a single core
single or multiple taxa - vs depth or time
### Plot of abundance of a single taxon in multiple cores
single or multiple taxa from multiple sites - vs depth or time
### Plot of abundance of a taxon vs an environmental characteristic 
(e.g., water chemistry variable)
-----
[[Provide more information on the functions - what they do.  get_sites and get_datasets are nearly the same. Can't use siteid or sitename with get_datasets. Get-downloads gets the data.  Use get_sites only if the only search info you have is site id or site name.]]

[[[We can often use a site search to do preliminary work since the `sites` object contains less information than the `datasets` or `downloads` objects. - It doesn't really seem to contain less information than a "get_datasets" object. DFC]]]