---
title: "Diatom Data Analysis"
author: "Don Charles, Eric O'Malley"
date: "11/18/2021"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup}
library(neotoma2)
library(dplyr)
library(purrr)
library(sf)
library(tidyverse)
library(leaflet)
library(leaflegend)
library(rioja)
library(readxl)
```

```{r setup1, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE)
```

#Data Manipulation

##Percent Abundance Function

The following code creates a percent abundance calculation of each sample from a sites object obtained from `get_downloads()`. A percent abundance column named `count_pct` is added to a dataframe containing other useful variables obtained from the `samples()` function.

```{r percent abundance function}
pc_abun <- function(dl_object) {
  
  samp <- samples(dl_object)
  age <- samp$age
  depth <- samp$depth
  taxa <- samp$variablename
  count <- samp$value
  sitename <- samp$sitename
  datasetid <- samp$datasetid
  siteid <- samp$siteid
  sampleid <- samp$sampleid
  lat <- samp$lat
  long <- samp$long
  
  
  countDF <- data.frame(lat, long, siteid, sitename, datasetid, sampleid, depth, age, taxa, count) %>%
    group_by(sampleid) %>%
    mutate(count_pct = count*100 / sum(count)) #THIS LINE PERFORMS THE CALCULATION
  return(countDF)
}
```



## Geo Political Units Function
(WIP)

The following code extracts the state using a site object
(change the [[2]] to [[3]] in the `geopol()` function to get county instead of state)

**Probably switch to searching by GPID in get_datasets, although this may still be useful**

```{r geo pol (state) function}
#x = sites object
geopol_form <- function(x) {
                allids <- getids(x)
                siteids <- unique(allids$siteid)
                return(siteids)
}

##USE OUTPUT FROM ABOVE FUNCTION INTO NEXT FUNCTION
##example: y <- geopol_form(x)
########## geopol(y)

geopol <- function(x) {
            states <- purrr::map(x, function(y) {
                                      site <- get_sites(y)
                                      dls <- get_downloads(site)
                                      dls[[1]]@geopolitical[[1]][[2]]
                                      })
            as.data.frame(x,(unlist(states))) %>%
            rename(siteid = x) %>%
            print()
}
```

```{r combine geopol() & samples()  function}
#function to add states to samples output dataframe
# x = geopol() output
# y = samples() output (or any dataframe with siteid)
sampdf <- function(x,y) {
x$states <- row.names(x)
  merge(y, x, by = 'siteid')
}
```

## Pivot Wider

Count data obtained from `samples()` is in a "long" format. Some common diatom analysis packages require the data be in a "wide" format (such as Rioja). The following code shows how to transform wider.

```{r, pivot wider}
tidyr::pivot_wider(widetable,
                             id_cols = age, #keeps longer
                             names_from = variablename, #column names
                             values_from = count_pct, #cell values
                             values_fill = 0)

```


#Aggregate Dataset Work

Aggregate dataset work involves using the excel file containing all of the aggregate dataset information. 

As of 6/10/2022: The file is named Aggregate Datasets - Diatoms -  EMAP NE SW Neotoma (version 1)

The file is located: \\files.drexel.edu\ANS\PCER_Data\Phycdata\PROJECTS\NEOTOMA\Aggreg Datasets - Dtm-WChem Corresp

The code chunks were initially made just copying and pasting the datasetids into R. 
**EDIT**
```{r agg dataset excel loading, eval=FALSE}
library(readxl)
# read_excel(path, sheet = NULL)
# Code needs to be modified to the path on your own computer with this file
## Can't seem to get it to work with file in Phycdata directory
aggregate_wWC <- read_excel("C:\\Users\\eco37\\Desktop\\R Files\\neotoma_diatom_wc\\neotoma_diatom_wc\\NE_Diat_Collab\\Aggregate Datasets - Diatoms -  EMAP NE SW Neotoma (version 1).xlsx", sheet = "Northeast combined wWC 28Apr22")

aggregate_wo_WC <- read_excel("C:\\Users\\eco37\\Desktop\\R Files\\neotoma_diatom_wc\\neotoma_diatom_wc\\NE_Diat_Collab\\Aggregate Datasets - Diatoms -  EMAP NE SW Neotoma (version 1).xlsx", sheet = "Northeast Combined wo WC 28Apr")
```


## RDS Object

Some of the sites objects are large. It may be worth saving as an RDS file to be read in to avoid downtime.

```{rds saving}
#saves any object to 
saveRDS(diat_taxa_tab, file = "diat_taxa_table.rds")
```


## States to Aggregate Dataset


```{R getting states from NLA agg dataset, eval=FALSE}
#NLA2007 = siteids from agg dataset
NLA2007 <- c(1222,1564,1766,2625,9594,9596,9716,10034,10091,10273,10365,11631,11666,11767,13061,13370,13393,13768,13840,14288,22680,22681,22682,22683,22684,22685,22686,22687,22688,22689,22690,22691,22692,22693,22694,22695,22696,22697,22698,22699,22700,22701,22702,22703,22704,23272,23273,23274,23275,23276,23277,23278,23279,23280,23281,23282,23283,23284,23285,23286,23287,23288,23289,23290,23295,23295,23295,23295,23295,23295,23295,23295,23295,23296,23297,23298,23299,23300,23301,23302,23303,23304,23305,23306,23307,23308,23309,23310,23311,23312,23313,23314,23315,23316,23317,23318,23319,23343,23344,23345,23346,23347,23348,23349,23350,23351,23352,23353,23354,23355,23356,23356,23356,23356,23357,23358,23359,23360,23361,23362,23363,23364,23365,23366,23367,23368,23369,23370,23371,23371,23371,23371,23372,23373,23374,23375,23376,23377,23378,23379,23380,23380,23380,23380,23381,23382,23383,23384,23384,23384,23384,23385,23386,24273,24276,24282,24283,24293,24294,24295,24318,24319,24331,24345,24355,24381,24385,24387,24388,24389,24390,24391,24392,24393,24394,24395,24396,24397,24398,24399,24400,24401,24402,24403,24404,24405,24406,24407,24408,24409,24410,24411,24412,24413,24414,24415,24416,24417,24418,24419,24420,24421,24422,24423,24424,24425,24426,24427,24428,24429,24499,24500,24501,24502,24504,24505,24506,24507,24508,24509,24510,24511,24512,24513,24514,24515,24516,24517,24518,24519,24520,24521,24522,24523,24524,24525,24526,24527,24528,24529,24530,24531,24532,24533,24534,24535,24536,24537,24538,24539,24540,24541,24542,24543,24544,24545,24546,24547,24548,24549,24550,24551,24552,24553,24554,24555,24556,24557,24558,24559,24560,24561,24562,24563,24564,24565,24566,24567,24568,24569,24570,24571,24572,24573,24574,24575,24576,24577,24578,24579,24580,24581,24582,24583,24584,24585,24586,24587,24588,24589,24590,24593,24594,24595,24596,24597,24598,24599,24600,24601,24602,24603,24604,24605,24606,24607,24608,24609,24610,24611,24612,24613,24614,24615,24616,24617,24618,24619,24620,24621,24622,24623,24624,24625,24626,24627,24628,24629,24630,24631,24632,24633,24634,24635,24636,24637,24638,24639,24640,24641,24642,24643,24644,24645,24646,24647,24648,24649,24650,24651,24652,24653,24654,24655,24656,24657,24658,24659,24660,24661,24662,24663,24664,24665,24666,24667,24668,24669,24670,24671,24672,24673,24674,24675,24676,24677,24678,24679,24680,24681,24682,24683,24684,24685,24686,24687,24688,24689,24690,24691,24692,24693,24694,24695,24696,24697,24698,24699,24700,24701,24702,24703,24704,24705,24706,24707,24708,24709,24710,24711,24712,24713,24714,24715,24716,24717,24718,24719,24720,24721,24722,24723,24724,24725,24726,24727,24728,24729,24730,24731,24732,24733,24734,24735,24736,24737,24738,24739,24740,24741,24742,24743,24744,24745,24746,24747,24748,24749,24750,24751,24752,24753,24754,24755,24756,24757,24758,24759,24760,24761,24762,24763,24764,24765,24766,24767,24768,24769,24771,24772,24773,24774,24775,24776,24777,24778,24779,24780,24805,24806,24807,24808,24810,24811,24812,24813,24814,24815,24816,24817,24819,24821,24823,24824,24826,24829,24830,24833,27895)


NLA_states2 <- purrr::map(NLA2007, purrr::possibly(function(x) {
  site <- get_sites(x)
  sitename <- site[[1]]$sitename
  geo <- site[[1]]@geopolitical[[1]] #change the geopolitical index number to "2" to get county instead
  list(sitename,geo)
}, otherwise = NA))

NLAstatedf <- t(as.data.frame(do.call(cbind, NLA_states2)))
colnames(NLAstatedf) <- c("Neotoma Site ID", "geopol")
NLAstatedf <- as.data.frame(NLAstatedf)
NLAstatedf <- (unnest(NLAstatedf_df, geopol))

new_NLAstatedf <- NLAstatedf_df %>%
  mutate(geopolchar=(as.character(NLAstatedf_df$geopol)), siteid = as.numeric(new_NLAstatedf$`Neotoma Site ID`))

new_NLAstatedf <- data.frame(new_NLAstatedf$geopolchar, new_NLAstatedf$siteid)
colnames(new_NLAstatedf) <- c("geopol", "Neotoma Site ID")

library(readxl)
NLA_2007_Excel <- read_excel("209 NLA 2007 Aggregate Dataset - REVISED 30 April 2022.xlsx") #Sheet: 209 NLA State 2007 - with states May2

final <- merge(new_NLAstatedf, NLA_2007_Excel, by = 'Neotoma Site ID')

#write.csv(final, file = "NLA_states_csv.csv")
```

## Leaflet Plot from Aggregate Dataset (Excel)

In the code chunk above we started with siteids from the aggregate dataset and ended with a dataframe that matched the states to the site information. Now, the code below filters for states of interest (NE states) and quickly plots them using the plotLeaflet() function.

```{r plotting NE_NLA}
#Agg #15
#Handle: NLA-2007
library(readxl)
NLA_states_excel <- read_excel("C:/Users/eco37/Desktop/NLA_states_excel.xlsx", 
               col_types = c("numeric", "text", "text", "text", "numeric", "text", "numeric","text", "text", "text", "text", "text", "text"))

NE_NLA <- dplyr::filter(NLA_states_excel, State %in% c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "New York", "New Jersey", "Rhode Island", "Vermont"))

NE_NLA_ids <- NE_NLA$`Neotoma Site ID`

NE_NLA_ids <- as.numeric(NE_NLA_ids)

NE_NLA_sites <- get_sites(NE_NLA_ids)

#using modified plotLeaflet, found locally in "Custom Functions"
NE_map <- plotLeaflet(NE_NLA_sites)

NE_map
```

# SCATTERPLOT
## Total Ph in WC against percent abundance

```{R plotting total Ph in WC against taxa % abundance}
#using: Aggregate Datasets - Diatoms -  EMAP NE SW Neotoma (version 1)
#sheet: EVERYTHING
#only using sites that have both a WC and DIAT datasetid
#ANSP / Drexel ID: 15
## commented out is example code of how to change for Phosphorus

DREXEL_agg <- read_excel("C:/Users/eco37/Desktop/Aggregate Datasets - Diatoms - for Neotoma 7-20-22.xlsx", sheet = "EVERYTHING", col_types = c("numeric", "numeric", "text","numeric","numeric"))

colnames(DREXEL_agg) <- c("drexelid","siteid","sitename","diatom_datasetid", "wc_datasetid") #changing column names to be more in line with neotoma output variables

with_wc_agg <- DREXEL_agg[!is.na(DREXEL_agg$wc_datasetid),] #RESULT: ONLY SITES THAT HAVE BOTH AN ASSOCIATED DIATOM AND WC DATASETS

#PLOTTING WATER CHEMISTRY #15
agg_wWC_15 <- with_wc_agg %>%
  dplyr::filter(drexelid == 15)

##Total Phosphorus (unfiltered)
##wc_diat_15_phos <- get_datasets(agg_wWC_15$wc_datasetid) %>%
  get_downloads() %>%
  samples() %>%
  dplyr::filter(variablename == "Total Phosphorus (unfiltered)")

#WC
wc_diat_15_ph <- get_datasets(agg_wWC_15$wc_datasetid) %>%
  get_downloads() %>%
  samples() %>%
  dplyr::filter(variablename == "pH") %>%
  dplyr::filter(elementtype == "field") %>%
  select(datasetid,siteid,sitename,units,value,variablename) %>%
  arrange(.,siteid)


#DIATOM
diat_15_abun <- 
  get_datasets(agg_wWC_15$diatom_datasetid) %>%
  get_downloads() %>%
  pc_abun()

diat_15_Eun <- diat_15_abun %>%
  dplyr::filter(taxa == "Eunotia exigua") %>%
  arrange(.,siteid)

wc_diat_15_ph_final <- subset(wc_diat_15_ph, wc_diat_15_ph$siteid %in% diat_15_Eun$siteid) #makes dataframe where only the wWC data for the sites where Eunotia exigua counts were found

x_ph <- wc_diat_15_ph_final$value
y_abun <- diat_15_Eun$count_pct

##phos_vs_abun <- plot(x_phos,y_abun,
   #xlab = "Total Phosphorus (µg/L)",
   #ylab = "% Abundance",
   #main = "Eunotia exigua vs. Phosphorus")

ph_vs_abun <- plot.default(x_ph,y_abun,
   xlab = "pH",
   ylab = "% Abundance",
   main = "Eunotia exigua vs. pH")

#adding linear regression line
reg = lm(y_abun ~ x_ph)
abline(reg, col = "red") #If there is the error "plot.net has not been called yet" then run plotting code in console


##phos_vs_abun <- plot(x_phos_no_190, y_abun_no_190, xlab = "Total Phosphorus (µg/L)", ylab = "% Abundance", main = "Eunotia exigua vs. Phosphorus")
```


#LEAFLET
## Adding Leaflet Markers to existing map **SHOULD BE CHANGED**
**NOT NEEDED** (can just use addCircleMarkers)
```{r plotting another dataset ontop of another function}
#data obtained from Diatom Aggregate Dataset
#ids = siteids
#previous map = leaflet object (results from plotLeaflet)
#circlecolor = hex color in quotes Ex:
add_dataset_leaflet <- function(ids, previousmap, circlecolor) {
  
  sites_ids <- get_sites(ids)
  
  form <- map(sites_ids@sites, function(x) {
    df <- data.frame(siteid = x@siteid,
                     sitename = x@sitename,
                     lat = mean(st_coordinates(x@geography)[, 2]),
                     long = mean(st_coordinates(x@geography)[, 1]),
                     elev = x@altitude,
                     description = x@description)
  }) %>%
    bind_rows()
  
  leaflet_map <- previousmap %>%
    addCircleMarkers(data = form, lng = form$long, lat = form$lat, radius = 0.5, color = circlecolor, opacity = 1,
                     popup = paste0("<b>", form$sitename,
                                    "</b><br><b>Description:</b> ",
                                    form$description,
                                    "<br><a href=http://apps.neotomadb.org/explorer/?siteids=",
                                    form$siteid,
                                    ">Explorer Link</a>"),
                     options = markerOptions(riseOnHover = TRUE))
  return(leaflet_map)
}
```

##Mapping diatom cores for NE

The results of the code below is a leaflet map with that plots taxa abundance with location. 

Makes use of a modified PlotLeaflet function in neotoma2 that I named PlotLeaflet2. This function is not in the document, but it is a separate script I have locally (add?). 

**NEED TO EDIT STILL**
```{R Agg Dataset Taxa Presence in Sites, warning=FALSE}
#note: this example includes states not in the NE. Shouldn't matter, since map is cropped to NE

#In the code above we got the same IDs using the Aggregate Excel sheet. This time, we will take all Diatom datasetids regardless if they have a corresponding WC datasetid.

agg_15 <- DREXEL_agg %>%
  dplyr::filter(drexelid == 15) 

diat_15_dl <- get_datasets(agg_15$diatom_datasetid) %>%
  get_downloads()

#############################
#this part will depend on what is the result of filtering for stelligera.
#so code cannot apply to all

samp_stelligera <- samples(diat_15_dl) %>%
filter(., grepl("stelligera", variablename, fixed = TRUE)) #This step is to see taxa with similar character strings. This can be useful in seeing if any taxa should be harmonized. For this example, we will not be harmonizing.

samp_stelligera <- samp_stelligera %>%
dplyr::filter(variablename == "Discostella stelligera") #filtering again to only have get Discostella stelligera

final <- samp_stelligera$siteid %>%
  get_sites() %>%
  plotLeaflet()
###################
# getting the sites where the taxa was not found

filtered_dataids <- samp_stelligera$datasetid

data_not_found <- setdiff(diat_15, filtered_dataids)
  
sites_not_found <- (as.data.frame(get_datasets(data_not_found)))$siteid
  
NLA_discostella_stelligera <- add_dataset_leaflet(sites_not_found, previousmap = final, circlecolor = '#999999')

######################## leaflet legend

NLA_discostella_stelligera <- NLA_discostella_stelligera %>% addLegend("bottomright", 
  colors =c("#03F", "#999999"),
  labels= c("Present","Not Present"),
  title= "D. stelligera distribution in NLA_2007",
  opacity = 1)


#result: leaflet map displaying which sites had atleast 1 count of the taxa vs. sites with none
```

```{R Agg Dataset Taxa Presence in Sites ALL}
#note: this includes states not in the NE. Shouldn't matter, since map is cropped to NE

#DATASETIDS OF AGG DATASET in 
# Included: diat_1, diat_4, diat_9, diat_15, diat_21, diat_22, diat_23, diat_24
########### diat_26, diat_32, diat_33, diat_34, diat_35, diat_36, diat_37  
diat_1 <- c(49022,49023,49024,49025,49026,49027,49028,49029,49030,49031,49032,49033,49034,49035,49036,49037,49038,49039,49040,49041,49042,49043,49044,49045,49046,49661,49662,49663,49664,49665,49666,49667,49668,49669,49670,49671,49672,49673,49674,49675,49676,49677,49678,49679,49680,49681,49682,49683,49684,49685,49688,49689,49690,49691,49692,49693,49694,49695,49696,49697,49698,49699,49700,49701,49702,49703,49704,49705,49706,49707,49708,49709,49710,49711,49712,49713,49714,49715,49716,49717,49718,49719,49720,49721,49722,49723,49724,49725,49726,49727,49728,49729,49730,49731,49732,49733,49734,49735,49736,49737,49738,49739,49740,49741,49742,49743,49744,49745,49746,49747,49748,49749,49750,49751,49752,49753,49754,49755,49756,49757,49758,49759,49760,49761,49762,49763,49764,49765,49766,49767,49768,49769,49770,49771,49772,49773,49774,49775,49776,49777,49778,49779,49780,49781,49782,49783,49784,49785,49786,49787,49788,49789,49790,49791,49792,49793,49794,49795,49796,49797,49798,49799,49800,49801,49802,49803,49804,49805,49806,49807,49808,49809,49810,49811,49812,49813,49814,49815,49816,49817,49818,49819,49820,49821,49822,49823,49824,49825,49826,49827,49828,49829,49830,49831,49832,49833,49834,49835,49836,49837,49838,49839,49840,49841,49842,49843,49844,49845,49846,49847,49848,49849,49850,49851,49852,49853,49854,49855,49856,49857,49858,49859,49860,49861,49862,49863,49864,49865,49866,49867,49868,49869,49870,49871,49872,49873,49874,49875,49876,49877,49878)

diat_1_taxa <- get_datasets(diat_1) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")


#############################################################################
diat_4 <- c(29964,30358,30735,30949,31878,31882,31958,40402,40403,40404,40455,40456,40457,40458,40459,40460,40461,40462,40463,40464,40465,40466,40467,40468,40469,40470,40471,40473,40474,40475,40476,40477,40478,40479,40480,43079,43132,43514,43515,43516,43518,43519,43526,43527,43528,43583,43584,43585,43586,43587,43588,43589,43590,43591,46236,49249,51982,51983,51984,51985,51986,51987,51988,51989,51990,51991,51992,51993,51994,51995,51996,51997,51998,51999,52000,52001,52002,52003,52004,52006,52007,52008,52009,52010,52011,52012,52013,52014,52015,52016,52017,52018)

diat_4_taxa <- get_datasets(diat_4) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")


#############################################################################
diat_9 <- c(44764,44765,44766,44767,44768,44769,44770,44771,44772,44773,44774,44775,44776,44777,44778,44779,44780,44781,44782,44783,44784,44785,44786,44787,44788,44789,44790,44791,44792,44793,44794,44795,44796,44797,44798,44799,44800,44801,44802,44803,44804,44805,44806,44807,44808,44809,44810,44811,44812,44813,44814,44815,44816,44817,44818,44819,44820,44821,44822,44823,44824,44825)

diat_9_taxa <- get_datasets(diat_9) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")

#############################################################################
diat_15 <- c(51593,51440,40058,40060,40063,40064,40075,40119,40135,40136,40138,40242,40244,40245,40248,40252,40268,40275,40278,40279,40283,51502,40286,51502,40286,40287,51595,51566,51681,51538,51452,50275,51660,51455,50417,51597,51497,40072,50248,50253,50255,50258,50267,50269,50273,50283,50285,50288,50289,50293,50379,50397,50411,50420,51443,51457,51463,51464,51467,51468,51469,51470,51475,51481,51495,51504,51524,51530,51533,51549,51550,51551,51554,51569,51570,51578,51586,51587,51598,51602,51605,51614,51617,51627,51638,51642,51648,51650,51654,51664,51669,51674,51675,51687,51689,51691,51694,50282,50284)

diat_15_taxa <- get_datasets(diat_15) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")

#percent_abun_taxa_map <- plotLeaflet2(diat_15_sites, df_with_count_pct = diat_15_taxa)

#############################################################################
diat_21 <- c(19321,19394,19400,19327,19329,19331,19333,19335,19337,19339,19402,19340,19404,19342,19344,19409,19346,19406,19411,19412,19415,19417)

diat_21_taxa <- get_datasets(diat_21) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")



#############################################################################
diat_22 <- c(19363,19364,19365,19366,19367,19368,19369,19370,19371,19372,19373,19374,19375,19376,19377,19382,19426,19431,19434,19437,19442,19445,19446,19447,19448,19449,19450,19451,19452,19453,19454,19455,19456,19457,19458,19459,19460,19461,19462,19463,19464,19473,19474,19477,19480,19483,19486,19489,19490,19491,19492,19505,19506,19507,19508,19509,19510,19511,19514,19517,19520,19523)

diat_22_taxa <- get_datasets(diat_22) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")

#############################################################################
diat_23 <- c(19872,19873,19876,19878,19880,19881,19882,19884,19886,19888,19890,19892,19894,19896)

diat_23_taxa <- get_datasets(diat_23) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")
#############################################################################
diat_24 <- c(19531,19535,19541,19545,19549,19555,19559,19563,19567,19571,19577,19579,19581,19594,19596,19598,19600,19602,19604,19606,19608,19610,19612,19614,19622,19624,19626,19628,19630,19632,19634,19636,19638,19640,19644,19650,19654,19656,19660,19664,19666,19668,19670,19672,19676,19678,19680,19682,19684)

diat_24_taxa <- get_datasets(diat_24) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")

#############################################################################
diat_26 <- c(19720,19722,19724,19726,19727,19729,19731,19733,19735,19736,19737,19738,19739,19740,19741,19742,19743,19745,19747,19749,19751,19752,19753,19754,19755,19757,19758,19759,19760,19761,19764,19767,19768,19769,19770,19771,19772,19773,19774,19775,19776,19777)

diat_26_taxa <- get_datasets(diat_26) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")

#############################################################################
diat_32 <- c(49957,49958,49959,49960,49961,49962,49963,49964,49967,49968,49969,49970,49971)

diat_32_taxa <- get_datasets(diat_32) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")


#############################################################################
diat_33 <- c(50077,50078,50079,50080,50081,50082,50083,50084,50085,50086,50087,50088,50089,50090,50091,50092,50093,50094,50095,50097,50098,50099,50100,50101,50102,50103,50104,50105,50106,50107,50108,50109,50110,50111,50112,50113)

diat_33_taxa <- get_datasets(diat_33) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")


#############################################################################
diat_34 <- c(50193,50194,50195,50196,50197,50198,50199,50200,50201,50202,50203,50204,50205,50206,50207,50208)

diat_34_taxa <- get_datasets(diat_34) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")


#############################################################################
diat_35 <- c(50213,50214,50209,50210,50211,50212)

diat_35_taxa <- get_datasets(diat_35) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")

#############################################################################
diat_36 <- c(50156,50157,50158,50159,50160,50161,50162,50163)

diat_36_taxa <- get_datasets(diat_36) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")


#############################################################################
diat_37 <- c(42748,42780,42844,42825,42798,42757,42749,42779,42828,42853,42796,42818,42812,42819,42777,42788,42754,42768,42799,42785,42759,42778,42787,42756,42753,42776,42850,42800,42843,42766,42822,42802,42817,42815,42811,42814,42807,42793,42755,42760,42835,42765,42792,42842,42772,42771,42840,42849,42761,42805,42804,42763,42762,42810,42784,42833,42809,42752,42808,42820,42783,42790,42782,42837,42797,42758,42827,42764,42795,42789,42767,42816,42829,42826,42839,42836,42750,42830,42848,42831,42847,42770,42834,42824,42794,42781,42823,42821,42786,42851,42846,42813,42806,42769,42775,42832,42791,42801,42852,42838,42841,42751,42803,42845)

diat_37_taxa <- get_datasets(diat_37) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Eunotia exigua")

##########################################################################################################################################################################################################################################################################################################################################################################################################

#combining all of the count data (defined above) from each agg dataset to one dataframe

all_sites <- rbind(diat_1_taxa, diat_4_taxa, diat_9_taxa, diat_15_taxa, diat_21_taxa, diat_22_taxa, diat_23_taxa, diat_24_taxa, diat_26_taxa, diat_32_taxa, diat_33_taxa, diat_34_taxa, diat_35_taxa, diat_36_taxa, diat_37_taxa)

all_taxa_formatted <- aggregate(all_sites$count_pct ~ all_sites$siteid + all_sites$lat + all_sites$long, data = all_sites, FUN = mean) %>%
  colnames() <- c("siteid", "lat", "long", "count_pct") %>%
  arrange() %>%
  get_sites(.$siteid)

croto_abun_taxa_map_3_1 <- plotLeaflet2(diat_sites_all_3_1, df_with_count_pct = alltaxa_3_test_1, title = "Eunotia exigua % Abundance")

croto_abun_taxa_map_3 <- plotLeaflet2(diat_sites_all_3, df_with_count_pct = taxa_table_all_3, title = "Eunotia exigua % Abundance")

```

## Visualization using Maps (plotLeaflet)


```{R abundance leaflet for one site (initial code)}
#Using custom function: plotLeaflet2
#only using NLA_2007
####this chunk (all the way to the below hashes) includes how to make a percent
####abundance taxa map
diat_15 <- c(51593,51440,40058,40060,40063,40064,40075,40119,40135,40136,40138,40242,40244,40245,40248,40252,40268,40275,40278,40279,40283,51502,40286,51502,40286,40287,51595,51566,51681,51538,51452,50275,51660,51455,50417,51597,51497,40072,50248,50253,50255,50258,50267,50269,50273,50283,50285,50288,50289,50293,50379,50397,50411,50420,51443,51457,51463,51464,51467,51468,51469,51470,51475,51481,51495,51504,51524,51530,51533,51549,51550,51551,51554,51569,51570,51578,51586,51587,51598,51602,51605,51614,51617,51627,51638,51642,51648,51650,51654,51664,51669,51674,51675,51687,51689,51691,51694,50282,50284)

diat_15_data <- get_datasets(diat_15)

diat_15_dl <- get_downloads(diat_15_data)

diat_15_pc <- pc_abun(diat_15_dl)

umm <- diat_15_pc %>%
dplyr::filter(taxa == "Discostella pseudostelligera")

umm_sites <- get_sites(umm$siteid)

percent_abun_taxa_map <- plotLeaflet2(umm_sites, df_with_count_pct = umm)

```




```{r plotting another dataset ontop of another 3}
#data obtained from Diatom Aggregate Dataset
#Agg #1
#Handle: EMAP_NE1

EMAP_NE1_ids <- c(26224,26225,26226,26227,26228,26229,26230,26231,26232,26233,26234,24385,23285,26235,26236,26237,26238,26239,26240,26241,26242,26243,26244,26245,26246,26247,26248,26249,26250,26251,26252,26253,26254,26255,26256,26257,26258,26259,26260,26261,24443,26262,26263,26264,26265,26266,23369,26267,13108,26268,23379,26269,23382,24289,26270,26271,26272,26273,13232,26274,26275,26276,26277,26278,26279,26280,26281,26282,26343,26344,26345,26346,25914,26347,26348,25916,25917,26349,26350,26351,26352,26353,26354,26355,26356,26357,26358,26359,26360,26361,26362,26363,26364,26365,26366,26367,26368,26369,26370,26371,26372,26373,26374,26375,26376,26377,26378,26379,26380,26381,26382,26383,26384,26385,26386,26387,26388,26389,26390,26391,26392,26393,22384,26394,26395,26396,26397,26398,26399,25935,26400,26401,26402,26403,26404,26405,26406,26407,26408,26409,26410,26411,26412,26413,26414,26415,26416,26417,26418,13105,26419,26420,26421,26422,26423,26424,26425,26426,26427,26428,26429,24649,26430,26431,26432,26433,26434,26435,26436,26437,26438,26439,26440,26441,13144,26442,26443,26444,26445,26446,26447,26448,26449,26450,26451,26452,26453,26454,26455,26456,26457,26458,26459,26460,26461,26462,26463,26464,26465,26466,26467,26468,26469,26470,26471,26472,24739,26473,26474,26475,26476,26477,26478,26479,24752,26480,26481,26482,26483,26484,26485,26486,26487,26488,24361,26489,26490,26491,26492,26493,26494,26495,26496,26497,26498,26499,26500,26501,26502,13056,26503)

NE_map <- add_dataset_leaflet(EMAP_NE1_ids, previousmap = NE_map, circlecolor = '#ff3333')
```

```{r plotting another dataset ontop of another 4}
#data obtained from Diatom Aggregate Dataset
#Agg #4
#Handle: ADK-CAL

ADK_CAL_ids <- c(17800,17990,720,14223,18648,18650,18688,710,711,712,23410,713,716,715,717,718,721,722,723,724,725,726,727,728,729,730,736,737,739,740,741,742,743,734,22911,17914,735,714,732,24460,744,746,24431,24432,24466,24433,24439,24442,24445,24458,24459,24463,24495,24435,24477,745,24436,24437,24438,24440,24443,24444,24446,24447,24456,24448,24451,24452,24453,24454,24455,24462,24464,24465,24467,24483,24482,24484,24486,24485,24489,24492,24494,13539,24497,24473,24475,24476,24478,24479,16574,24481)

NE_map <- add_dataset_leaflet(ADK_CAL_ids, previousmap = NE_map, circlecolor = '#ffff00')

```

```{r plotting another dataset ontop of another 5}
#data obtained from Diatom Aggregate Dataset
#Agg #9
#Handle: NENGPIRL

NENGPIRL_ids <- c(25899,25900,25901,25902,25903,13048,25904,24185,25905,25906,25907,25908,25909,25910,25911,25912,18865,25913,25914,25915,25916,25917,25918,25919,25920,25921,25922,25923,25924,25925,25926,25927,25928,18872,25929,25930,25931,1615,25932,25933,25934,25935,25936,25937,25938,25939,25940,25941,25942,25943,25944,25945,25946,25947,25948,25949,25950,25951,25952,25953,25954,25955,25956)

NE_map <- add_dataset_leaflet(NENGPIRL_ids, previousmap = NE_map, circlecolor = '#00cc00')
```

```{r plotting another dataset ontop of another 6}
#data obtained from Diatom Aggregate Dataset
#Agg #24, 26 (25 missing from combined ?)
#Handle: NJ-PAL10, NJ_PAL14

NJ_PAL10_ids <- c(13158,13160,13163,13165,13167,13170,13172,13174,13176,13178,13181,13182,13183,13190,13191,13192,13193,13194,13195,13196,13197,13198,13199,13200,13201,13202,13203,13204,13205,13206,13207,13208,13209,13210,13211,13212,13213,13214,13215,13216,13217,13218,13219,13220,13222,13223,13224,13225,13226)

NJ_PAL14_ids <- c(13241,13232,13242,13178,13243,13244,13202,13203,13209,13245,13246,13215,13247,13219,13223,13224,13225,13248,13249,13250,13251,13252,13253,13254,13255,13256,13257,13258,953,13259,13261,13263,13264,13265,13266,13267,13268,13269,13270,13271,13272,13273)

NJ_ids <- c(NJ_PAL10_ids,NJ_PAL14_ids)

NE_map <- add_dataset_leaflet(NJ_ids, previousmap = NE_map, circlecolor = '#9900ff')
```

```{r plotting another dataset ontop of another 7}
#data obtained from Diatom Aggregate Dataset
#Agg #23
#Handle: GG-PONDS
GGreener_ids <- c(13317,13318,13319,13320,13321,13322,13323,13324,13325,13326,13327,13328,13329,13330)

NE_map <- add_dataset_leaflet(GGreener_ids, previousmap = NE_map, circlecolor = '#ff9900')

```

```{r plotting another dataset ontop of another 8}
#data obtained from Diatom Aggregate Dataset
#Agg #32, 33, 34, 35, 36
#Handle: VT-REF11, VT-NLA12, VT-LA13, VT-1W14, VT-OL18
VT_ids <- c(8587,24372,24369,24363,24362,24360,24368,24367,24366,24370,24365,24371,24364,25901,13132,27816,27817,27818,24369,24417,27819,27820,13049,27821,27822,27823,13142,27824,27825,27829,26467,27826,27827,24417,27828,27796,27797,27798,27799,24583,27800,26449,27801,26224,27716,27717,27718,27719,27720,27721,27722,26274,24417,27723,27724,27725,13049,27726,24355,27727,27728,27729,27730,27731,24537,27732,24356,27733,13580,27734,27735,23152,27736,27737,24706,27738,27739,27740,27741)

NE_map <- add_dataset_leaflet(VT_ids, previousmap = NE_map, circlecolor = '#000000')

```


```{r plotting another dataset ontop of another 9}
#data obtained from Diatom Aggregate Dataset
#Agg #21
#Handle: NJ-NY-PL

NJ_NY_PL_ids <- c(13027,13067,13028,13029,13030,13031,13032,13033,13034,13035,13070,13036,13071,13037,13038,13074,13039,13072,13075,13076,13078,13080)

NE_map <- add_dataset_leaflet(NJ_NY_PL_ids, previousmap = NE_map, circlecolor = '#00ffff')

```

```{r plotting another dataset ontop of another 10}
#data obtained from Diatom Aggregate Dataset
#Agg #22
#Handle: NE-PALEO

NE_PALEO_ids <- c(13045,13046,13047,13048,13049,13050,13051,13052,1688,13053,13054,13055,13056,13057,13058,13061,13084,13087,1697,13090,13093,13095,13096,13097,13098,13099,13100,13101,13102,13103,13104,13105,13106,13107,13108,13109,13110,13111,13112,13113,13114,13119,13120,13122,13124,13126,13128,13130,13131,13132,13133,13140,13141,13142,13143,13144,13145,13146,13148,13150,13152,13154)

NE_map <- add_dataset_leaflet(NE_PALEO_ids, previousmap = NE_map, circlecolor = '#663300')

```


### Leaflet Legend
```{r leaflet legend}
#data obtained from Diatom Aggregate Dataset

NE_map <- NE_map %>% addLegend("bottomright", 
  #colors used in circle markers above
  colors =c("#03F", "#ff3333", "#ffff00", "#00cc00", "#9900ff", '#ff9900', '#000000', '#00ffff', '#663300'),
  labels= c("NLA-2007","EMAP_NE1","ADK-CALIBR","NENG-PIRLA","NJ-PALEO", "GG-PONDS", "VT-PALEO", "NJ-NY-PALEO", "NE-PALEO"),
  title= "NE Diatom Surface Samples",
  opacity = 1)

```











# STRATIGRAPHIC PLOTTING
## GGPLOT spatial-based data (GGPLOT, RIOJA)

```{r GGPLOT spatial-based data}
#code obtained from Simon Goring's R_Workshop neotoma2 presentation (May 15, 2022)
alldiatsurf <- get_datasets(datasettype="diatom surface sample")
surface_dl <- get_downloads(alldiatsurf)
surfacesamp <- samples(surface_dl)

#Using Spatial-based Data (July max temperature)
spatial <- sf::st_as_sf(surfacesamp, 
                        coords = c("long", "lat"),
                        crs = "+proj=longlat +datum=WGS84")

worldTmin <- raster::getData('worldclim', var = 'tmax', res = 10)
surfacesamp$tmin7 <- raster::extract(worldTmin, spatial)[,7]

#choosing taxa
minsamp <- surfacesamp %>% distinct(siteid, .keep_all = TRUE) %>% 
  select(tmin7)

topten <- surfacesamp %>% 
  group_by(variablename) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))

surfsubsamp <- surfacesamp %>% filter(variablename %in% topten$variablename[1:10])

#plot
library(ggplot2)

ggplot() +
  geom_density(data = surfsubsamp,
               aes(x = round(tmin7 / 10, 0)), col = 2) +
  facet_wrap(~variablename) +
  geom_density(data = minsamp, aes(x = tmin7 / 10)) +
  xlab("Minimum July Temperature") +
  ylab("Kernel Density")
```

```{r RIOJA epd_binder strat example, warning=FALSE}
#code modified from https://open.neotomadb.org/EPD_binder/simple_workflow.html
diatom_dl <- get_datasets(datasettype = "diatom", limit = 1) %>%
  get_downloads()

#This line takes the first site from the object, in this case we only had
#one site in the object anyways, so its not needed
plottingSite <- diatom_dl[[1]]

#I don't think this is needed. Output of taxa only contains 1 ecologicalgroup and elementtype (which is DIAT & valve, respectfully)
#CODE DOESN'T WORK FOR ME, SKIPPED
plottingTaxa <- taxa(plottingSite) %>%
  filter(ecologicalgroup %in% c("DIAT")) %>%
  filter(elementtype == "valve") %>%
  arrange(desc(samples)) %>% 
  head(n = 10)

#I dont think we need to filter ???
#SKIPPED
shortSamples <- samples(plottingSite) %>% 
  filter(variablename %in% plottingTaxa$variablename) %>% 
  filter(ecologicalgroup %in% c("DIAT")) %>%
  filter(elementtype == "valve") %>%
  filter(units == "NISP")

#code that skips the filtering process but retrieves taxa and samples output
plottingTaxa <- neotoma2::taxa(plottingSite) %>%
  head(n = 10)

shortSamples <- samples(plottingSite) %>% 
  filter(variablename %in% plottingTaxa$variablename)

#unsure of what "proportion value" is, will try and just use percent abundance.
onesite <- shortSamples %>%
  group_by(age) %>%
  mutate(diatomcount = sum(value, na.rm = TRUE)) %>%
  group_by(variablename) %>% 
  mutate(abundance = value*100 / diatomcount) %>% 
  arrange(desc(age))


#
widetable <- onesite %>%
  dplyr::select(age, variablename, abundance) %>% 
  mutate(abundance = as.numeric(abundance))

counts <- tidyr::pivot_wider(widetable,
                             id_cols = age,
                             names_from = variablename,
                             values_from = abundance,
                             values_fill = 0)


#plotting attempt
clust <- rioja::chclust(dist(sqrt(counts)),
                        method = "coniss")

plot <- rioja::strat.plot(counts[,-1], yvar = counts$age,
                  title = diatom_dl[[1]]$sitename,
                  ylabel = "Calibrated Years BP",
                  xlabel = "Diatom (%)",
                  y.rev = TRUE,
                  clust = clust,
                  wa.order = "topleft", scale.percent = TRUE)

rioja::addClustZone(plot, clust, 4, col = "red")
```


stratigraphic diagram of depth (y) vs. abundance (x) of taxa above a specified abundance in **ONE site**

```{r GGPLOT rforpaleo strat ONE SITE example}
# produces stratigraphic diagram of depth (y) vs. abundance (x) of taxa above a specified abundance in ONE site.

###################################################################DATA PREP
diatom_cores <- get_datasets(datasettype = "diatom", limit = 1) %>%
  get_downloads() %>%
  samples()

diatom_df <- data.frame(diatom_cores$depth, diatom_cores$age, diatom_cores$variablename, diatom_cores$value)

colnames(diatom_df) <- c("depth", "age", "taxa", "count")

diatom_abun <- diatom_df %>%
  group_by(depth) %>%
  mutate(relative_abundance_percent = count / sum(count) * 100)

diatom_common_taxa <- diatom_abun %>%
  group_by(taxa) %>%
  summarise(max_rel_abund = max(relative_abundance_percent)) %>%
  filter(max_rel_abund >= 10) %>%
  arrange(max_rel_abund) %>%
  pull(taxa)

diatom_counts_common <- diatom_abun %>%
  filter(taxa %in% diatom_common_taxa) %>%
  mutate(taxa = factor(taxa, levels = diatom_common_taxa)) %>%
  arrange(taxa)


######################################################################PLOT

species_plot_obj <- ggplot(
  diatom_counts_common, 
  aes(y = depth, x = relative_abundance_percent)
) +
  # draw horizontal lines of the appropriate length for each depth
  geom_segment(aes(xend = 0, yend = depth), lwd = 1) +
  # facet by taxon, keeping distance on the x axis comparable
  # between facets
  facet_grid(~taxa, scales = "free_x", space = "free_x") +
  # reverse the y axis for depth
  scale_y_reverse() +
  # make all facets use the same break values
  # (helps with cluttered breaks)
  scale_x_continuous(breaks = c(0,2,4,6,8)) +
  # set the x and y labels
  labs(x = "Relative Abundance (%)", y = "Depth (cm)") +
  # customize the appearance
  theme(
    # rotate the facet labels
    strip.text.x = element_text(angle = 60, hjust = 0, vjust = 0), 
    # turn off the label background
    strip.background = element_blank()
  )

# voodoo that makes it so that facet labels can overlap
# https://stackoverflow.com/questions/49740215/ggplot-facet-grid-label-cut-off
species_plot_grob <- ggplotGrob(species_plot_obj)
for(i in which(grepl("strip-t", species_plot_grob$layout$name))){
  species_plot_grob$grobs[[i]]$layout$clip <- "off"
}

# needed to draw the modified plot_grob
grid::grid.draw(species_plot_grob)
```


## MULTIPLE sites stratigraphic diagram of age (y) vs. abundance (x) of a single taxa.

```{r GGPLOT rforpaleo strat MULTIPLE SITE example}
# goal: produce stratigraphic diagram of age (y) vs. abundance (x) of a single taxa of MULTIPLE sites.

###################################################################DATA PREP

#takes first 5 sites listed as diatom datasettype in Neotoma
um <- get_datasets(datasettype = "diatom", limit = 5) %>%
  get_downloads() %>%
  samples()

um2 <- data.frame(um$depth, um$age, um$variablename, um$value) #selecting variables

colnames(um2) <- c("depth", "age", "taxa", "count") #changing column names

um3 <- um2 %>%
  group_by(age) %>%
  mutate(relative_abundance_percent = count / sum(count) * 100)

um_common_taxa <- um3 %>%
  group_by(taxa) %>%
  summarise(max_rel_abund = max(relative_abundance_percent)) %>%
  filter(max_rel_abund >= 2) %>%
  arrange(max_rel_abund) %>%
  pull(taxa)

um_counts_common <- um3 %>%
  filter(taxa %in% um_common_taxa) %>%
  mutate(taxa = factor(taxa, levels = um_common_taxa)) %>%
  arrange(taxa)


get_datasets(46241)
get_datasets(46242)
get_datasets(46233)

one <- get_datasets(46241) %>%
  get_downloads() %>%
  samples()

two <- get_datasets(46242) %>%
  get_downloads() %>%
  samples()

three <- get_datasets(46233) %>%
  get_downloads() %>%
  samples()


x1 <- unique(one$variablename)
x2 <- unique(two$variablename)
x3 <- unique(three$variablename)
Reduce(intersect, list(x1, x2, x3))

one_1 <- one %>%
  dplyr::filter(variablename == "Melosira ambigua")

two_1 <- two %>%
  dplyr::filter(variablename == "Melosira ambigua")

three_1 <- three %>%
  dplyr::filter(variablename == "Melosira ambigua")


all <- c(46241,46242,46233)

all_ambigua <- get_datasets(all) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Melosira ambigua")


species_plot_ob <- ggplot(
  all_ambigua, 
  aes(y = age, x = count_pct)
) +
  # draw horizontal lines of the appropriate length for each depth
  geom_segment(aes(xend = 0, yend = age), lwd = 1) +
  # facet by taxon, keeping distance on the x axis comparable
  # between facets
  facet_grid(~sitename, scales = "free_x", space = "free_x") +
  # reverse the y axis for depth
  scale_y_reverse() +
  # make all facets use the same break values
  # (helps with cluttered breaks)
  scale_x_continuous(breaks = c(0, 1, 5, 15, 25)) +
  # set the x and y labels
  labs(x = "Relative Abundance (%)", y = "age (years before 1950)", title = "Melosira ambigua") +
  theme_bw() +
  # customize the appearance
  theme(
    # rotate the facet labels
    strip.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0),
    #center title
    plot.title = element_text(hjust = 0.5),
    # turn off the label background
    strip.background = element_blank()
  )

species_plot_ob
```


#Diatoms and Water Chemistry
Aggregate Dataset, getting count data for diatom and corresponding water chemistry
```{r diatom water chemistry agg dataset example}


```

