---
title: "Working with diatom data in the neotoma2 R-package - June 2022"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{The neotoma2 R Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
```

```{r setup, message=FALSE}
library(neotoma2)
library(dplyr)
library(purrr)
library(sf)
library(tidyverse)
```

# Introduction

This document provides general guidance and specific code examples for those wanting to use `neotoma2` to retrieve and work with diatom data from the Neotoma Paleoecology Database. This vignette, as part of the `neotoma2` package, also provides the package developers (Socorro Dominguez and Simon Goring) a set of structured descriptions for the kinds of work that diatom researchers may wish to perform with Neotoma data and some of the specific tasks they may wish to accomplish. Researchers who would like to add examples or other workflows can reach out to the developers through [GitHub Issues](https://github.com/NeotomaDB/neotoma2/issues), or the [Neotoma Slack Workspace](https://join.slack.com/t/neotomadb/shared_invite/zt-cvsv53ep-wjGeCTkq7IhP6eUNA9NxYQ).

A main goal of `neotoma2` is to pull records from the Neotoma database and into a user's R workflow. This will help diatomists and other environmental scientists produce clear working documents that use data from Neotoma and allow them to combine it with their own data or data from other sources.

This document is organized to address users' two main goals: to browse the database to discover datasets that meet specified criteria and to create output files that can be used for further analysis. Other goals are to create plots and maps and analyze data using R code.

### Data organization in Neotoma

It is worth knowing how the Neotoma database is organized and what we mean when talking about things like `sites`, `datasets`, and `samples`.

The Neotoma data model is available online at the [Database Schema](https://open.neotomadb.org/dbschema) page. As of September 2021, there were 150 different tables in the database. Some of these have particular kinds of data (for example, `agetypes`), and some link tables together. The `neotoma2` package directly supports access to some tables, but not all. For specialized cases, a user may be required to access an entire table directly (using the `get_table()` function). Neotoma is a Postgres database, and people who are comfortable with SQL can download a snapshot of the database and run much of the analysis on their own.

#### Neotoma Data Structure

Neotoma data is structured spatially. Each collection location is assigned a **site**, a spatial location defined by either a point or a polygon. It may be a lake, an archaeological site, a swamp, or another location.

Within a **site**, there may be one or more **collection units**.  
A collection unit represents the material (e.g., sediment core) removed from a specific location within a **site** (e.g., a lake). So, a researcher may have taken multiple cores within a lake basin, or an archaeologist may have several dig sites within a larger area. Collection unit material is subdivided into **analysis units** (e.g., sediment intervals within a core). **Samples** are defined as material removed from analysis units and analyzed for a particular dataset type (e.g., diatoms, pollen, water chemistry).   Groups of analysis units (samples) from an individual collection unit are then grouped into a **dataset**.

Note: There are three diatom dataset types:  "diatom" (stratigraphic / cores), "diatom surface samples" (surface samples only), and "diatom top-bottom" (top and bottom only).

#### neotoma2 R-package functions

The `neotoma2` package uses data from the Neotoma database, but it is mediated through R functions (like `get_sites()` to get site information) that call public URLs, which are abstractions of database calls. So, for example, to get data for the site with ID number 12:
```r
get_sites(12)
```
calls out to the internet for the result at:
```
https://api.neotomadb.org/v2.0/data/sites/12
```

Which itself calls an SQL query:
```sql
SELECT * FROM ndb.sites
WHERE siteid = 12
```

It is a lot easier for most users to use the R functions since they create objects in R that work with other commands and statistical analysis they might want to do.

There are three main functions in `neotoma2` to retrieve and subset data: `get_sites()`, `get_datasets()`, and `get_downloads()`.  Each returns a `sites` object output with increasing level of detail (e.g., `get_sites` least; `get_downloads` most).  A `sites` object can potentially contain a complete set of data for one or more sites, including information on site locations, collection units, chronologies, datasets samples and taxa counts.   
Neotoma2 also includes helper functions such as `filter()`, `collunits()`, `datasets()`, `samples()`, and `plotLeaflet()`.  They are used to extract elements from the obtained `sites` objects.  

Functions are described where they first appear in the document below.  See the `neotoma2` package documentation for more details.


# Browse Data

This section describes ways of searching the Neotoma database to find `sites` and diatom `datasets` using various criteria. The primary purpose of each code chunk is to find the `sites` or `datasets` that meet the criteria specified in the code. The returned objects can then be used as input to other code chunks that return more detailed information on the datasets and create output files for data analysis (e.g., diatom counts).

### Find sites

Use the `get_sites()` or `get_datasets()` functions to find sites and related information. Searches using siteid or sitename can be done using `get_sites()`. Searches using dataset information (such as age, location, dataset type, etc.) can be done using `get_datasets()`. 

The information returned by both functions is similar, but still notably different. The level of detail in `get_datasets()` is different as it includes: agerange, database, datasettype, and datasetnotes. These fields are not in `get_sites()` output.  [[TRUE??]]

Both functions can be used to search for information on a single or multiple sites. They are helpful for finding locations of all sites within a region, even if they are not diatom sites; this is useful for indicating the level of paleoecological 'effort' in a region.  Also, output can be used to generate a map of all sites and superimpose specific sites of interest.

#### `get_sites` function

Find sites using the `get_sites` function.

**Usage:** Sites can be searched using the following supported parameters
(e.g., `get_sites(sitename="Alexander Lake")`).

| parameter | allowed values |
| ------ | ------- |
| `siteid` | integers (values > 0) |
| `sitename` | valid character string; `%` as wildcard |
| `altmax` | integers (values >= than `altmin`) |
| `altmin` | integers |
| `loc` | spatial coordinates (WKT or geoJSON) |
| `limit` | integer > 0; default 25 |
| `offset` | integer > 0; default 0 |

`limit` indicates the maximum number of records that will be returned; `offset` indicates the number in a sequence where processing will begin.  Together, these allow large numbers of records to be processed in batches.

**Values returned:** A `sites` object, which includes nested lists with information on: siteid, sitename, latitude and longitude, elevation, geopolitical unit, notes, site description, collection units, and datasets associated with the collection units.  In addition to the data above, the function returns a table (dataframe) with 5 location variables for each site (see example below).

##### Search by site ID

Get data for a single site using siteid. 

```{R dearch by siteid}
uwallf_site <- get_sites(745) # finds site data for Upper Wallface Pond, which has the Neotoma siteid 745.

uwallf_site
```
To see the structure of the output, type "str(uwallf_site)" in the RStudio console.

Get data on multiple sites by using siteids.

Example input data:
| `sitename` |  `siteid` |
| -- | -- |
| Jackson Pond | 25918 |
| Jellison Hill Pond | 25919 |
| Jimmie Pond | 25920 |

Get data on multiple sites by using siteids
```{r multsitesbySiteIDs example}
siteids <- c(25918, 25919, 25920)
diat_6 <- get_sites(siteids)
diat_6
```
Get data on multiple sites by reading a text file of siteids

The above code returns data for three sites.  We could use a larger set of site IDs either by loading the IDs from a file, or by entering them directly.  Here we assume that there exists a plain text file called `NLAD.txt` with a list of site IDs for sites sampled as part of the National Lakes Assessment.  The contents of the file look like this [[(the ellipsis would be replaced by other numbers)]]:

There may well be many more lines with numbers representing each site identifer in the set of records.  In this case we would load the data like so:

```{r NLAmultiplesiteID}
nlad <- read.table('NLAD.txt', header=FALSE)
nlaSites <- get_sites(nlad$V1)
nlaSites
```


##### Search by site name

If we know the name of one or more of the existing sites we are looking for, we can get site data using the `sitename` parameter.

Get data on a single site using sitename.

```{r sitesbyName example}
# diat <- get_sites(sitename = "")  # Basic format
diat <- get_sites(sitename = 'Schrader Pond')  # Example
diat
# wildcards (%) can be used if you do not know the exact name
# The`sitename` input is not case sensitive 
## Example: get_sites(sitename = '%Schra%')
```


Get data on multiple sites using sitenames.

```{r multsitesbySitename example}
# Join a list of site names together using the combine command, c()
sitenameList <- c("Bald Hill", "Blake (sutton)", 
                  "High (sudbry)", "Holland", 
                  "Little Hosmer", "Long (grnsbo)", 
                  "Long (shefld)", "Round (shefld)")
my_sites <- c()
for(sitename in sitenameList){
  a_site <- get_sites(sitename = sitename)
  my_sites <- c(my_sites, a_site)
}
my_sites
```

##### Search for sites in a latitude and longitude bounding box

Example of finding sites by location within a bounding box of lats & longs.  Same code works with `get_datasets()` as well as `get_sites()`.
[[Revise wording in following code chunk; clarify use of the two BB options]]
```{R diatomsbyBoundingBox}
# Find sites in the northeast US.  Order of boundaries is: West, South, East, North.  
northeastUS <- rbind(c(-80,40), c(-80,48),
                     c(-65,48), c(-65,40), c(-80,40))
neus <- sf::st_polygon(list(northeastUS))

# The above code also works with: 
# northeastUS <- c(-80, -65, 48, 40)

#As the default, `get_datasets()` only retrieves the first 25 results. Use the argument `limit =" to change this.
neUSsites <- get_datasets(loc = neus, limit = 10)

neUSsites
```

#### Search for sites in geopolitical units

Find all sites located in a geopolitical unit as specified by a gpid. There are up to four hierarchical levels of geopolitical unit (first to fourth) to which a site can be assigned.  The combination of names of levels for a site depends on the country or region in which a site is located.  In the USA the four levels are country, state, county and administrative unit (e.g., Jeferson National Forest).

[[A gpid is a useful way to specify country and state boundaries. - delete?]]
The code below will open the Geopolitical units table.  Once the table appears, use the search function at top of the page to find the name of a geopolitical unit of interest and its corresponding gpid. 

```{R geopoliticalunitsTable}
loc <- neotoma::get_table(table.name ='geopoliticalunits')
View(loc)
# Use the filter at the top of the page to find name of a geopolitical unit and its corresponding gpid
```

```{r get site by gpid}
# diat_1<-get_site(gpid = x) # basic format
diat_1<-get_sites(gpid = 6198) # Example using geopolitical ID for Alaska - GeoPoliticalUnits table.  Returns table of sites located in Alaska. 
diat_1
```
View the sitegeopolitical table to find gpids associated with datasetids.

Run following code to open the sitegeopolitical table and then enter datasetid in filter box to find corresponding gpids

```{R sitegeopoliticalTable}
dataset_gpid <- get_table(table.name ="sitegeopolitical")
View(dataset_gpid)
# Table columns: GeoPoliticalID,  SiteID, SiteGeoPoliticalID
```

Example - get sites from multiple geopolitical units (e.g., states)

```{r retrieveByGeopol}
# Get site data for all sites in three New England states (ME, VT, NH, CT, MA, RI) plus NJ and NY [[DFC - delete some states from list above- include only the 3 states used]]
# The gpids were obtained from the Geopolitical units table 

gpids <- c(8412, 7990, 7934) #gpids for .......

ne_sites <- c()
for (id in gpids) {
  a_site <- get_sites(gpid = id)
  ne_sites <- c(ne_sites, a_site)
}
#returns data for first 25 sites
ne_sites
```

##### Navigate the structure of a `sites` object to find specific metadata.

Sometimes it is useful to extract specific data from a `sites` object.  We can call any individual site using `[[ ]]`, placing the index (sequence number) of the desired site object between the brackets. For example, the `uwallf_site` object is composed of only one site, so its index value is 1. Then we can also call the particular variable we want (e.g., siteid) using the `$` symbol.

```{r extractElement}
#Create a `sites` object using get_sites; this example - Upper Wallface Pond
uwallf_site <-get_sites(sitename = 'Upper Wallface%')
#To retrieve a variable (siteid, for example) from the first item (site) in the list, put a "1" in the brackets followed by "$siteid".
uwallf_site[[1]]$siteid
```

### Find datasets

Find datasets using the `get_datasets` function

##### `get_datasets` function

Use this function to discover datasets available for a geographic location, within an elevation range, representing a range of ages, and more.  Also use it to retrieve metadata for specific datasets.

**Usage:** `get_datasets(x, sites_object, contactid, datasettype, altmin, altmax, loc, ageyoung, ageold, ageof)`

x is an optional value that represents an object with numeric datasetids. It can be a single datasetid, an array with multiple IDs, or a `sites` object created using `get_sites`.

**Values returned:** A `sites` object with more detail than a `get_sites` object that includes information on collection unitsa nd datasets (e.g., datasettype, dataset notes and dataset age range).

##### Search using datasetid

Get data for a single dataset using datasetid 

```{r get datasetid}
# diat_ID <- get_dataset(datasetid) - basic format
diat_ID <- get_datasets(44764) # Example - (The) Horns Pond diatom dataset
diat_ID
```

Get data on multiple datasets using multiple datasetids 
```{r multdatsetid }
multid <- c(44764,44765,44766)
diat_multid <- get_datasets(multid) # Example
diat_multid
```

#### Search using elevation range

Find datasets for sites between minimum and maximum elevations (m)

```{R getDatasetssByElevExample}
# Find datasets for sites ranging in elevation from 500 to 5500 m
diat_el <- get_datasets(altmin=500,altmax=5500, limit = 10)
# Returns data for the first 25 sites
diat_el
```

###### Structure of `get_datasets` output

To see structure of a `get_datasets()` object, run the following code:

```{R structure_of_get_dataset_output}
uwallf_site <-get_sites(sitename = 'Upper Wallface%') # finds site data for a site, including datasetid's.
uwallf_data <- get_datasets(uwallf_site) # finds datasets for Upper Wallface Pond; returns 4 elements corresponding to the 4 datasets available for U Wallface Pond.  Each element is a different datasettype:  diatom surface sample, diatom, chronology and water chemistry.

#str(uwallf_data)
```

# Download Data
## Downloads function

Download all records associated with one or more datasets. Use the `get_downloads` function to return a `sites` object with full level of detail.

**Usage:**  `get_downloads(x)`
Arguments: `x` can be a datasetid or a vector of numeric datasetids. It can also be a `sites` object as returned by `get_datasets()` or `get_sites()`; "verbose, logical (default is TRUE)" - prints a download status bar.

**Values returned:** a `sites` object with a full level of detail, including data on sites, collection units, analysis units, samples, datasets, taxonomic data (e.g., diatom counts), chronologies, and contacts.

To see the structure of `get_downloads()` function output, run the following code:
```{R structure_of_get_download_output}
uwallf_site <-get_sites(sitename = 'Upper Wallface%') # finds site data for a site
uwallf_data <- get_datasets(uwallf_site) # finds datasets data for Upper Wallface (there's 4 elements)
uwallf_diatom_dataset <- neotoma2::filter(uwallf_data, datasettype == "diatom") # only keep diatom datasets
uwallf_download <- get_downloads(uwallf_diatom_dataset)

#to see a quick summary of collectionunits, chronologies, datasets, and datasetttypes, use the `summary()` function
summary(uwallf_download)

#str (uwallf_download) # remove "#" from beginning of this line and run code to show complete structure of uwallf_download
```

## Helper functions

After obtaining a `get_downloads` result, a typical next step is to create subsets of data for intended analyses.  The following helper functions can be used on a `sites` object to return a subset of associated values as a data frame without having to go through the structure of lists returned from `get_sites`, `get_datasets`, or `get_downloads`.

Some examples of helper functions are `datasets()`, `collunits()`, `samples()`,`taxa()`, `chronologies()` and `coordinates()`. These are listed and described within the `neotoma2` package documentation. Below are some basic examples of the output of the first three of the functions listed above.

#### `datasets()` function
The `datasets()` function extracts dataset information from a `sites` object. 
**Usage:** `datasets(object)`

The argument for this function is a `sites` object.

**Values returned:** A datasets object with the variables dataset.id, database, datasettype, age_range_old, age_range_young, and notes.


```{R datasetsFunctionExample}
#search for a site by site name
uwallf_site <- get_sites(sitename = 'Upper Wallface%')
#get datasets from get_sites output
uwallf_data <- get_datasets(uwallf_site)
#filter the datasets output to include only "diatom" datasets
diat_uwallf_dataset <- neotoma2::filter(uwallf_data, datasettype == "diatom")
#use the datasets function to get the data output
diat_uwallf_data <- datasets(diat_uwallf_dataset)
#view the data frame of the data returned
diat_uwallf_data

#to save the output in current working directory
write.csv(diat_uwallf_data, row.names = FALSE)
# Example of how a csv file could be saved to a directory other than the working directory:
# write.csv(diat_uwallf_data,"C:\\Users\\eco37\\Desktop\\datasets_uwallf.csv", row.names = FALSE)
```

##### Create dataframe of dataset information

The `datasets()` function returns an object with basic dataset information. The following code is an example of how to convert that object to a dataframe.

```{R dataframeDatasetsoutput}
as.data.frame(datasets(uwallf_data))

#as a dataframe, the variables (columns) can be called
#Example:
(as.data.frame(datasets(uwallf_data)))$datasetid
```

The functions used to extract elements from `sites` objects (like `datasets()`) may not return all the variables of interest, or too many variables. Custom tables, or data frames, can be made from the information within the `sites` object. The following example creates a table containing the variables obtained from the `datasets()` function, along with additional variables like siteid and sitename.

```{R custom_datasets_dataframe}
uwallf_site <- get_sites(sitename = 'Upper Wallface%')
uwallf_data <- get_datasets(uwallf_site)
uwallf_dl <- get_downloads(uwallf_data)

#SITE DATASETS - defining variables to be put into data frame
#This code (specifically for SITEID and SITENAME) only works for one site)
SITEID <- rep(uwallf_dl[[1]]@siteid,(length(as.data.frame(datasets(uwallf_data))$datasetid)))
SITENAME <- rep(uwallf_dl[[1]]@sitename,(length(as.data.frame(datasets(uwallf_data))$datasetid)))
DATASETID <- as.data.frame(datasets(uwallf_data))$datasetid
#DATASETNAME <-
DATASETTYPE <- as.data.frame(datasets(uwallf_data))$datasettype
AGE_RANGE_OLD <- as.data.frame(datasets(uwallf_data))$age_range_old
AGE_RANGE_YOUNG <- as.data.frame(datasets(uwallf_data))$age_range_young
#FIRST PUBLICATION <-

custom_datasets <- data.frame(SITEID,SITENAME,DATASETID,DATASETTYPE,AGE_RANGE_OLD,AGE_RANGE_YOUNG)

custom_datasets

#Save the output as a CSV file, which can be opened in Microsoft Excel
write.csv(custom_datasets, row.names = FALSE)
```

#### `collunits()` Function

The `collunits` function extracts collection units from a `sites` object. 

**Usage:** `collunits(object)`
The argument for this function takes a "sites" object.

**Values returned:** A datasets object with the variables dataset.id, database, datasettype, age_range_old, age_range_young, and notes.

```{R collunitsFunctionExample}
#search for a site by site name
uwallf_site <- get_sites(sitename = 'Upper Wallface%')
#get datasets from get_sites output
uwallf_data <- get_datasets(uwallf_site)
#filter the datasets output to include only diatoms
diat_uwallf_dataset <- neotoma2::filter(uwallf_data, datasettype == "diatom")
#take the diatom datasets and getting the downloads output
uwallf_dwl <- get_downloads(diat_uwallf_dataset)
#using the samples function to get the data from the collunits function
diat_uwallf_collunits <- collunits(uwallf_dwl)
#view the collection unit data
diat_uwallf_collunits

#Save the output as a CSV file, which can be opened in Microsoft Excel.
#The file will be saved in the working directory.
write.csv(diat_uwallf_collunits, row.names = FALSE)
```

#### `samples()` Function
The `samples()` function obtains sample level data from a record `sites` object and returns results as a data frame. 

**Usage:** `samples(x)`

**argument:** (x) is a `sites` object obtained from `get_downloads`.

**Values returned:** A data frame with the variables age, agetype, ageolder, ageyounger, chronologyid, chronologyname, units, value, context, element, taxonid, symmetry, taxongroup, elementtype, variablename (taxa), ecologicalgroup, depth, thickness, sample name, datasetid, siteid, sitename, lat, long, area, sitenotes, description, elev, collunitid, database, datasettype, age_range_old, age_range_young, and datasetnotes.

##### Diatom counts

The following code chunks can be used to view diatom count related data.  There are a few examples, each with something a little different.
For single sites and multiple sites; aggregate data sets; stratigraphic vs surface sample vs top-bottom; percentages vs raw counts; containing specific taxa, etc.

Example of using `samples()` to obtain diatom counts from a single sediment core.

```{R samplesFunctionExample}
# Join a list of site names together using the combine command, c()
  sitenameList <- c("Bald Hill", "Blake (sutton)", 
                    "High (sudbry)", "Holland", 
                    "Little Hosmer", "Long (grnsbo)", 
                    "Long (shefld)", "Round (shefld)")
my_sites <- c()
for(sitename in sitenameList){
  a_site <- get_sites(sitename = sitename)
  my_sites <- c(my_sites, a_site)
}
my_sites

#get datasets using the get_sites output
NE_datasets <- get_datasets(my_sites)
#use the get_datasets output to get downloads output
diat_NE_dl <- get_downloads(NE_datasets)

#filter the get_downloads output to include only diatoms
#result: diatom count information for the specified sites
NE_diat_samp <- samples(diat_NE_dl) %>%
dplyr::filter(datasettype == "diatom surface sample")

#output is a large table. The head function is used to show only the first three rows to save space
head(NE_diat_samp, 3)

#Example code to save the output as a CSV file, [[(which can be opened in Microsoft Excel)]] in a location other than the working directory:
#write.csv(diat_samp_uwallf,"C:\\Users\\eco37\\Desktop\\diat_uwallf.csv", row.names = FALSE)
```

##### Water chemistry

Example of using `samples()` to obtain water chemistry data for multiple lakes.

```{r waterchemistry_counts example}
# Use the same sites specified in the chunk above. An object of the site data has already been made.
my_sites

NE_datasets <- get_datasets(my_sites)

wat_NE_dataset <- neotoma2::filter(NE_datasets, datasettype == "water chemistry")

wat_NE_dl <- get_downloads(wat_NE_dataset)  

#result: count information with only water chemistry at the specified sites. 
head(samples(wat_NE_dl),3)
```

##### Diatom counts and water chemistry
Example of how to get sample data from multiple sites while also filtering for diatom and water chemistry datasets. 

The `sites` objects obtained from the above two chunks of code can be combined to have both diatoms and water chemistry data.

```{r diatom_and_waterchemistry_counts example, warning=FALSE}
#Join downloads together using the combine command, c()
diat_wat_dl <- c(diat_NE_dl, wat_NE_dl)

#result: count information with both diatoms and water chemistry at the specified sites
diat_wat_counts <- samples(diat_wat_dl)
```

#### `chronologies()` Function
The `chronologies()` function obtains chronology data from a record(s) and returns results as a list. 

**Usage:** `chronologies(x)`

**argument:** (x) is a `sites` object (that contains chronologies) obtained from `get_downloads`.

**Values returned:** A list with the variables chronologyid, notes, contact, agemodel, ageboundolder, ageboundyounger, isdefault (?), dateprepared, modelagetype, chronologyname, and chroncontrols.

```{R chronologies, eval=FALSE}
uwallf_dl <- get_sites(sitename = 'Upper Wallface%') %>%
  get_datasets() %>%
  get_downloads()
 
uwallf_chron <- chronologies(uwallf_dl)
```


##### Diatom taxa

The following code gets selected diatom taxa name variables from the Neotoma Taxa table.

```{R MasterListTaxa}
#By default, get_table only returns 25 specimen records
diat_taxa_tab <- get_table('taxa', limit=99999) %>%
  dplyr::filter(taxagroupid == "DIA")

head(diat_taxa_tab, 5)
```


Get list of all taxa associated with single datasettype.

```{R ListOfTaxaByDatasettype, warning=FALSE}
diatoms<- get_datasets(datasettype = 'diatom')
diat_dl <- get_downloads(diatoms)
diat_taxa <- taxa(diat_dl) # List of taxa in all datasets of dataset type "diatom"

head(diat_taxa,3)
```

##### `get_publications` Function

The `get_publications` function extracts collection units from a `sites` object. 

**Usage:** `get_publications(x = NA, ...)`

| parameter | allowed values |
| ------ | ------- |
| `x` | integer: a contact ID |
| `	publicationid ` | The unique numeric identifier associated with a publication in Neotoma |
| `datasetid` | A unique identifier for a Neotoma dataset that is associated with a publication |
| `siteid` | A unique identifier for a Neotoma site that is associated with a publication |
| `pubtype ` | The publication type, from `get_tables("publicationtypes")` |
| `year` | The year the publication was released |
| `search` | A plain text search string used to search the citation |

**Values returned:** A "publications" object that displays the variables publicationid, citation, and DOI. The structure of the "publications" object contains metadata with all possible publications info.

```{R publications example}
uwallf_site <- get_sites(sitename = 'Upper Wallface%')

uwallf_data <- get_datasets(uwallf_site)

#filtering for only diatom datasets
diat_uwallf_dataset <- neotoma2::filter(uwallf_data, datasettype == "diatom")

#using `getids()` to retrieve the diatom datasetids
diat_uwallf_datasetids <- getids(diat_uwallf_dataset)$datasetid 

#`get_publications()` can only take one datasetid at a time. Putting "[1]" after the object takes the first datasetid.
uwallf_pub <- get_publications(datasetid = diat_uwallf_datasetids[1])

uwallf_pub
```

# Visualize Data

Make plots and maps 
(e.g., stratigraphic diagrams, maps of taxa distributions)

## Map of distribution of sites

## `plotLeaflet()` function

The `neotoma2` package has a function to map a `sites` object. Use it, for example, to create a map showing all sites that are located within a specific geographic region, are part of an aggregate dataset, etc. 

```{R siteSpatialSearch1}
northeastUS <- rbind(c(-80,40), c(-80,48),
                     c(-65,48), c(-65,40), c(-80,40))
neus <- sf::st_polygon(list(northeastUS))

#As the default, `get_datasets()` only retrieves the first 25 results. Use the argument `limit =" to change this.
neUSsites <- get_datasets(loc = neus, limit = 199)

my_map <- plotLeaflet(neUSsites)

my_map
``` 

The map returned by `plotLeaflet()` is a `leaflet` object. Leaflet items (as defined in the `leaflet` package) can then be added to the map as desired.

```{R siteSpatialSearch2}
library(leaflet)
my_map %>% addCircleMarkers(lng =-80, lat = 45, color = "red") #adds a red marker
``` 
